---
title: BEAT CF MARS
subtitle: Simulation Report
description: |
    Three-arm Bayesian adaptive trial nested within BEAT-CF platform evaluating antibiotic-sparing strategies for managing pulmonary exacerbations
date: last-modified
date-format: "D MMMM YYYY"
author: 
  - name: Mark Jones
    id: mj
    email: mark.jones1@sydney.edu.au
version: 0.1
sponsor: University of Newcastle, NSW, Australia
protocol-number: Version 1.1  01AUG2024
registration: todo
hrec: todo
ci1: Tom Snelling
editor: source
bibliography: ../etc/refs.bib
csl: ../etc/elsevier-harvard.csl
# number-sections required otherwise section refs will not render 
number-sections: true
toc: true
toc-depth: 3
format:
  pdf: 
    pdf-engine: xelatex
    keep-tex: true
    documentclass: scrreprt
    papersize: a4
    fontsize: 12pt
    mainfont: Libertinus Serif
    sansfont: Libertinus Sans
    monofont: Libertinus Mono
    mathfont: Libertinus Math
    linestretch: 1.25
    template-partials: 
      - "../_extensions/partials/before-body.tex"
    include-in-header:
      text: |
       \usepackage{physics}
       \setkomafont{chapter}{\fontsize{16}{18}\selectfont}
       \setkomafont{section}{\fontsize{14}{16}\selectfont}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r}
#| echo: false

# uml digs
suppressPackageStartupMessages(library(nomnoml))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(qs))
suppressPackageStartupMessages(library(git2r))
suppressPackageStartupMessages(suppressWarnings(library(gt)))
suppressPackageStartupMessages(library(ggh4x))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))

toks <- unlist(tstrsplit(getwd(), "/")) 
if(toks[length(toks)] == "beatcf-mars-sim"){
  prefix_cfg <- "./etc/sim01/"
  prefix_stan <- "./stan"
  prefix_fig <- "./fig"
  prefix_data <- "./data"
} else {
  prefix_cfg <- "../etc/sim01/"
  prefix_stan <- "../stan"
  prefix_fig <- "../fig"
  prefix_data <- "../data"
}

# Each input file corresponds to the results from a single simulation
# scenario/configuration.
# Load all the files into a single list.

# files of interest
sim_lab <- "sim01-01"

flist <- list.files(paste0(prefix_data, "/", sim_lab), pattern = "sim01")
toks <- list()
l <- list()
i <- 1

for(i in 1:length(flist)){

  l[[i]] <- qs::qread(file.path(paste0(prefix_data, "/", sim_lab), flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

N_sims <- l[[1]]$cfg$nsim
```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  BEAT-CF MARS |
|Reference intervention: |  Standard airway clearance therapy at least twice per day for 14 days, plus 14 days of oral antibiotics commencing as soon as practicable (and ≥ 72 hours and < 7 days) after onset of new or acute worsening of cough. |   
|Intervention: | (1) Deferred antibiotic strategy (2) Early discontinuation antibiotic strategy. |   
|Outcome: | Parent/self-report of any of: (1) worsened respiratory symptoms on Day 4 compared to Day 0, (2) no improvement in symptoms on Day 7, (3) failure to return to pre-morbid baseline symptoms on Day 14, or (4) breathlessness impacting normal activity, or otherwise need to default to clinician-directed care at any time on or before Day 14 |   
|Study design:  |   Bayesian adaptive trial with early stopping rules | 
|Sponsor:  |    todo | 
Protocol: |  todo |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Tom Snelling | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}

| Version    |   Date     | Change    |   Reason     |
|:----|:------------|:----|:------------|
| 0.1 | 2025-07-11 | First version | N/A |


{{< pagebreak >}}



# Introduction

This report documents the simulation approach and results for the operating characteristics for the BEAT-CF MARS study.
The report is an operational document that will be updated, as necessary, over the course of the study.
It should be read in conjunction with the relevant version of the statistical analysis plan.

We provide the data generation assumptions, modelling approaches, scenarios and results that were used to explore the design.

 These results are based on simulation ``r sim_lab`` with ``r N_sims`` simulated trials run per scenario.

# Data generation

Data is generated based on subject matter expertise and while necessarily a simplification of reality, it aims to capture the aspects that are essential to the design.
The distributional assumptions of each data component follows.

We simulate design variables for lung function (ppFEV1 either >= or <70%) and pseudomonas airway colonisation (no pseudo or pseudo) in the 24 months prior to enrolment and treatment assignment.

Lung function is based on a multinomial distribution with two levels with the probability of having good lung function (ppFEV1 >= 70%) being 60%.
Conditional on lung function, pseudomonas colonisation is simulated based on a multinomial distribution with two levels.
The probability of pseudomonas colonisation given lung function >= 70% is set to 33% and and the probability of pseudomonas colonisation given lung function < 70%  is set to 0.63%.
The above values were arbitrarily selected.

Within each lung function/colonisation strata, treatment status was allocated 1:1:1.

Occurrence of treatment failure was simulated as a bernoulli random variable with probability formed from a linear risk model with a single outcome generated for each participant.

The participant characteristics and their outcome variables are generated at the start of each interim analysis so that the data accrues sequentially for the analyses.
As the trial progresses, decisions may be made which lead to early stopping of treatment arms.

To speed up parameter estimation, we aggregate the number of successes and number of trials by covariate group so that we can adopt an analogous binomial likelihood.

# Modelling {#sec-modelling}

## Overview

While the data are simulated using a linear risk model, the simulation model is specified as a multivariable logistic regression model from which we subsequently transform the parameters back to the risk scale via the inverse link and a g-computation step.
The model form is:

$$
\begin{aligned}
y &\sim \text{Binomial}(\pi, n) \\
\text{logit}(\pi)  &=  \alpha + \beta_{[\text{lung}]} + \gamma_{[\text{pseudo}]} +  \delta_{[\text{trt}]}  \\
\end{aligned}
$$

where $y$ is a binomial variable for the number of events out of $n$ trials for a distinct covariate pattern occurring with probability $\pi$ calculated from the linear predictor as follows:

+ $\alpha$ reference level log-odds of a successful outcome
+	$\beta_j$ effect of lung function (< 70% pFEV1, >= 70% pFEV1)
+	$\gamma_k$ effect of pseudomonas colonisation (not colonised, colonised)
+	$\delta_l$ effect of treatment (soc, deferred, early discontinuation)

In practice, the log-odds of treatment failure may be differential for locality status within each region, but this has been ignored for now and no interactions are included.

The model uses priors:

+ $\alpha \sim \text{Logistic}(-0.84, 0.4)$
+ $\beta_j \sim \text{Normal}(0, 3)$
+ $\gamma_k \sim \text{Normal}(0, 3)$
+ $\delta_l \sim \text{Normal}(0, 3)$

all of which are on the log-odds or log-odds ratio scale.
Transformed to the probability scale, the intercept prior is shown in @fig-prior-intercept and reflects the prior probability of treatment stratergy failure in the reference covariate groups.

```{r, echo = F, eval = T}
#| label: fig-prior-intercept
#| fig-cap: 'Prior on intercept'
#| fig-height: 4
#| fig-width: 4
#| fig-pos: H

d_fig <- data.table(
  x = rlogis(1e6, -0.84, 0.4)
)

ggplot(d_fig, aes(x = plogis(x))) +
  geom_density() +
  scale_x_continuous("Intercept", breaks = seq(0, 1, by = 0.1)) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.box="horizontal",
    strip.text.y.right = element_text(angle = 0,
                                      hjust = 0,
                                      vjust = 0.2,
                                      size = 7),
    axis.ticks = element_blank(),
    strip.text.x.top = element_text(size = 7),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey",
                                  linewidth = 0.1,
                                  linetype = 1),
    axis.title.y=element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.text.x =  element_text(size = 7),
    axis.title.x = element_text(size = 7),
    legend.text = element_text(size = 7)
  ) 
```


## Alternatives

Various alternative modelling perspectives might be considered:

1. Obtain an ordinal representation of the primary outcome
2. Consider re-randomisation

Given that non-inferiority is the main focus of the study, assuming a dose-response perspective could be anti-conservative and is therefore not advised.

# Decision procedures

The decision processes are based on the current data and are interpreted with reference to the current available evidence.
This approach is simple and offers a transparent interpretation, but it ignores the possibility that accumulating data may shift the posterior.
The decision thresholds and evidential thresholds are shown in @tbl-dec-thresholds.

```{r}
#| echo: FALSE
#| label: tbl-dec-thresholds
#| tbl-pos: H
#| tbl-cap: "Decision threshold parameters"



d_dec_pars <- data.table(
  desc = c("NI", "Inferiority"),
  ref_value = c(l[[1]]$cfg$dec_delta_ni, l[[1]]$cfg$dec_delta_inf),
  threshold = c(l[[1]]$cfg$dec_thresh_ni, l[[1]]$cfg$dec_thresh_inf)
)


gt_tbl <- gt(d_dec_pars) |>
  cols_align(
    columns = 1,
    align = "left"
  ) |>
  cols_align(
    columns = 2:3,
    align = "center"
  )  |>
  cols_label(
    desc = "Decision type",
    ref_value = "Reference value",
    threshold = "Threshold"
  ) |>
  tab_options(
    table.width = pct(60),
    table.font.size = pct(65)) 

gt_tbl 

```

In the current implementation, we evaluate non-inferiority and inferiority at each interim and, if a decision threshold is met, then we will stop recruitment into the relevant arm.
This approach is adopted for each interim and the final analysis and is reflected in the current set of results.

An alternative setup might consider reference arm switching when an intervention arm is indicated as non-inferior.
For example, assuming the Deferred stratergy was identified as non-inferior, we might halt enrolment into the SoC arm and continue to evaluate the Discontinue strategy relative to the Deferred strategy, which would replace the original SoC and act as a new reference arm.
Under this decision structure, care is needed to avoid degradation of treatment effects (aka biocreep/efficacy drift) and therefore subsequent evaluations might switch to evaluate superiority rather than non-inferiority. 
That is, in order for the Discontinue strategy to be recommended as a treatment policy, it would need to be shown to be superior to the Deferred stratergy.

The trial setup is such that if the intervention were beneficial, it would reduce the failure rate and we are therefore considering positive non-inferiority margins.
Specifically, if an intervention arm is non-inferior to the reference arm, then the failure rate on the intervention will be elevated by no more than the non-inferiority margin.
If the probability that the failure rate on the intervention arm is lower than the non-inferiority margin is very high, then we conclude non-inferiority.
Contrastingly, for the inferiority rule, if the probability that the risk difference is greater than the non-inferiority margin is high, then we will conclude inferiority.

+ For non-inferiority, we evaluate $\text{Pr}(RD < \epsilon_{NI}) > \zeta_{NI}$ where $\epsilon_{NI}$ and $\zeta_{NI}$ are pre-specified and correspond to a non-inferiority margin and an evidential requirement in terms of a probability threshold.
+ For inferiority, we evaluate $\text{Pr}(RD > \epsilon_{inf}) > \zeta_{inf}$ where $\epsilon_{inf}$ and $\zeta_{inf}$ are pre-specified correspond to a reference value (currently we set $\epsilon_{inf} = \epsilon_{NI}$) and an evidential minimum in terms of a probability threshold.





# Scenarios

Each scenario adopts a maximum sample size of 600 with interim analyses run after 300 enrolments have reached their primary endpoint and every 100 thereafter.
Given the use of a linear risk model in the data generation process, the treatment effects were specified as risk differences.

All scenarios used fixed covariate distributions and effects over the duration of the study. 
Additionally, all simulations used the same reference values and decision thresholds.

```{r, echo = F, eval = T}
#| label: scenario_list
#| code-summary: Scenarios

# Cumulative probability of decisions:

# Traverse the list of simulation results and for each one summarise the 
# cumulative probability of each decision type.

i <- 1
d_scenarios <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  l_cfg <- copy(l[[i]]$cfg)
  
  btrt <- unlist(l_cfg$btrt)
  
  d_scenarios <- rbind(
    d_scenarios,
    data.table(
      id = i,
      desc = l_cfg$desc,
      rd_2_1 = btrt[2] - btrt[1],
      rd_3_1 = btrt[3] - btrt[1]
    )
  )
}


```

```{r, eval = T}
#| echo: FALSE
#| label: tbl-scenarios
#| tbl-pos: H
#| tbl-cap: "Simulation scenarios"

gt_tbl <- gt(d_scenarios) |>
  cols_width(
    id ~ pct(10),
    desc ~ pct(50),
    rd_2_1 ~ pct(20),
    rd_3_1 ~ pct(20)
  ) |>
  cols_align(
    columns = 1:2,
    align = "left"
  ) |>
  cols_align(
    columns = 3:4,
    align = "center"
  )  |>
  tab_spanner(
    label = md("Effect size (risk difference) relative to SoC"),
    columns = 3:4
  ) |>
  cols_label(
    id = "ID",
    desc = "Scenario",
    rd_2_1 = "Deferred",
    rd_3_1 = "Discontinue"
  ) |>
  tab_options(
    table.width = pct(90),
    container.width = pct(90),
    table.font.size = pct(65)) 

gt_tbl 
```


```{r, echo = F, eval = T}
#| label: risk-grid
#| code-summary: Strata-level risk

# Cumulative probability of decisions:

# Traverse the list of simulation results and for each one summarise the 
# cumulative probability of each decision type.

i <- 1
d_grid_all <- data.table()

d_grid <- CJ(
  lung = 1:2,
  pseudo = 1:2,
  trt = 1:3,
  rd_2_1 = NA_real_,
  rd_3_1 = NA_real_
)



# For each scenario that was simulated
for(i in 1:length(l)){
  l_cfg <- copy(l[[i]]$cfg)
  
  alpha <- l_cfg$bmu
  b_lung <- unlist(l_cfg$blung)
  b_pseudo <- unlist(l_cfg$bpseudo)
  b_trt <- unlist(l_cfg$btrt)
  
  d_grid[, p := alpha + b_lung[lung] + b_pseudo[pseudo] + b_trt[trt]]
  
  d_grid[, rd_2_1 := b_trt[2] - b_trt[1]]
  d_grid[, rd_3_1 := b_trt[3] - b_trt[1]]
  
  d_grid_all <- rbind(
    d_grid_all,
    data.table(
      id = i,
      desc = l_cfg$desc,
      rd_2_1 = d_grid$rd_2_1,
      rd_3_1 = d_grid$rd_3_1,
      lung = d_grid$lung,
      pseudo = d_grid$pseudo,
      trt = d_grid$trt,
      p = d_grid$p
    )
  )
}


```


```{r, echo = F, eval = T}
#| label: fig-strata-risk
#| fig-cap: 'Underlying (true) probability of treatment failure by strata and scenario'
#| fig-height: 7
#| fig-width: 6
#| fig-pos: H



d_fig <- copy(d_grid_all)
d_fig[, trt := factor(trt, levels = 1:3, labels = c("Soc", "Deferred", "Discontinue"))]
# d_tbl_1_cur <- dcast(d_tbl_1_cur, id + desc ~ reg + loc + trt, value.var = "p")

# d_fig[, desc := sprintf("%s (RD = %.3f)", desc, rd_trt)]

d_fig[, lung := factor(lung, levels = 1:2, labels = c("<70%", ">=70%"))]
d_fig[, pseudo := factor(pseudo, levels = 1:2, labels = c("no pseudo", "pseudo"))]

ggplot(d_fig, aes(x = trt, y = p)) +
  geom_col(aes(fill = trt), 
           width = 0.15, lwd = 0.7) +
  # scale_color_discrete("") +
  scale_linetype_discrete("") +
  scale_fill_manual("", values = c("#E66100", "#5D3A9B", "black")) +  
  scale_x_discrete("") +
  scale_y_continuous("Pr(treatment failure)", breaks = seq(0, 0.5, by = 0.1)) + 
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.box="horizontal",
    strip.text.y.right = element_text(angle = 0,
                                      hjust = 0,
                                      vjust = 0.2,
                                      size = 7),
    axis.ticks = element_blank(),
    strip.text.x.top = element_text(size = 7),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey",
                                  linewidth = 0.1,
                                  linetype = 1),
    axis.title.y=element_text(size = 7),
    axis.text.y = element_text(size = 7),
    axis.text.x =  element_blank(),
    # axis.text.x = element_text(size = 7, angle = 90, vjust = 0, hjust=0.4),
    axis.title.x = element_text(size = 7),
    legend.text = element_text(size = 7)
  ) +
  facet_grid2(
    desc ~ lung + pseudo,
    labeller = label_wrap_gen(20)) 
```


# Results


## Probability of triggering decisions

@tbl-cprob-decision provides the cumulative probability of superiority by scenario with the probability of declaring futility in parentheses.
@fig-cprob-decision gives a visual representation of the same data.

```{r, echo = F, eval = T}
#| label: cum_prob_dec
#| code-summary: Cumulative probability of each decision type

# Cumulative probability of decisions:

# Traverse the list of simulation results and for each one summarise the 
# cumulative probability of each decision type.

i <- 1
d_cprob_dec <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_pr_dec[par %in% c("rd_2_1", "rd_3_1")])
  # config for scenario
  l_cfg <- copy(l[[i]]$cfg)
  
  # number of enrolments at each interim (interim sample size sequence)
  d_N <- data.table(ia = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  d_dec_1[, cdec := as.integer(cumsum(dec)>0), keyby = .(sim, rule, par)]
  d_dec_1[, cdec := nafill(cdec, type = "locf"), keyby = .(sim, rule, par)]
  d_dec_1[, cdec := as.logical(cdec)]
  
  d_dec_1 <- merge(d_dec_1, d_N, by = "ia")
  # cumulative proportion for which each decision quantity has been met by 
  # analysis and domain
  d_dec_1 <- d_dec_1[, .(pr_val = mean(cdec)), keyby = .(ia, N, rule, par)]
  
  
  d_cprob_dec <- rbind(
    d_cprob_dec,
    cbind(scenario = i, desc = l_cfg$desc, d_dec_1)
  )

}
```


```{r, eval = T}
#| echo: FALSE
#| label: tbl-cprob-decision
#| tbl-cap: 'Cumulative probability of decision at each interim'
#| tbl-pos: H



d_tbl_1_cur <- copy(d_cprob_dec)
d_tbl_1_cur <- dcast(
  d_tbl_1_cur, scenario + desc + par ~ rule + N, value.var = "pr_val")

d_tbl_1_cur[par == "rd_2_1", par := "Deferred"]
d_tbl_1_cur[par == "rd_3_1", par := "Discontinued"]


d_b_trt <- data.table()

for(i in 1:length(l)){
  l_cfg <- copy(l[[i]]$cfg)
  btrt <- unlist(l_cfg$btrt)
  
  d_tmp <- data.table(scenario = i, par = c("Deferred", "Discontinued"), rd = c(btrt[2]-btrt[1], btrt[3]-btrt[1]))
  d_b_trt <- rbind(d_b_trt, d_tmp)
}

d_tbl_1_cur <- merge(d_tbl_1_cur, d_b_trt, by = c("scenario", "par"))


setcolorder(d_tbl_1_cur, c("scenario", "desc", "par", "rd"))
setorderv(d_tbl_1_cur, cols = c("scenario"))

d_tbl_1_cur <- d_tbl_1_cur[, .SD, .SDcols = !c("scenario")]

g_tbl <- d_tbl_1_cur |> 
  gt(groupname_col = "desc") |> 
  gt::text_transform(
    locations = cells_row_groups(),
    fn = function(x) {
      lapply(x, function(x) {
        gt::md(paste0("*", x, "*"))
      })
    }
  ) |>
  cols_align(
    columns = 1:2,
    align = "left"
  )  |> 
  cols_align(
    columns = 3:ncol(d_tbl_1_cur),
    align = "center"
  )  |> 
  cols_merge(
    columns = c("ni_300", "inf_300"
                ),
    pattern = "<<{1}>><< ({2})>>"
  )   |> 
  cols_merge(
    columns = c("ni_400", "inf_400"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |> 
  cols_merge(
    columns = c("ni_500", "inf_500"
                ),
    pattern = "<<{1}>><< ({2})>>"
  )  |> 
  cols_merge(
    columns = c("ni_600", "inf_600"
                ),
    pattern = "<<{1}>><< ({2})>>"
  )   |>
  cols_width(
    starts_with("ni") ~ px(90)
  ) |>
  tab_spanner(
    label = md("Participants having reached primary endpoint"),
    columns = 4:ncol(d_tbl_1_cur)
  )  |>
  cols_label(
    rd = "Risk diff relative to SoC",
    ni_300 = html("300"),
    ni_400 = html("400"),
    ni_500 = html("500"),
    ni_600 = html("600")
  ) |>
  tab_options(
    table.font.size = pct(65),
    latex.use_longtable = TRUE,
    latex.header_repeat = TRUE 
  ) |>
  fmt_number(decimals = 2, drop_trailing_zeros = F)

g_tbl
```


{{< pagebreak >}}

```{r, echo = F, eval = T}
#| label: fig-cprob-decision
#| fig-cap: 'Cumulative probability of decision at each interim (enrolment by interim)'
#| fig-height: 7
#| fig-width: 7
#| fig-pos: H


d_fig_1 <- copy(d_cprob_dec)

d_fig_1 <- merge(
  d_fig_1, 
  melt(d_scenarios[, .(id, rd_2_1, rd_3_1)], id.vars = "id", value.name = "rd") ,
  by.x = c("scenario", "par"),
  by.y = c("id", "variable")
)

d_fig_1[, rule := factor(
  rule, levels = c("ni", "inf"), labels = c("NI", "Inferiority"))]


d_fig_1[, par := factor(
  par, levels = c("rd_2_1", "rd_3_1"), labels = c("Deferred", "Discontinued"))]

# d_fig_1[, desc := factor(
#   desc, 
#   levels = c(
#     "Null scenario", 
#     "Treatment reduces chance of RSV-LRI",
#     "Treatment increases chance of RSV-LRI"
#     )
# )]

d_fig_1[, group := paste0(par, " - ", rule)]

ggplot(d_fig_1, aes(x = N, y = pr_val, group = group , 
                    col = par, lty =  rule)) +
  geom_line(lwd = 0.25) +
  # scale_color_discrete("") +
  scale_linetype_manual("Decision type", 
                        values = c("solid", "dashed")) +
  scale_color_manual("Treatment stratergy", 
                     values = c("#E66100", "#5D3A9B")) +
  scale_x_continuous("Enrolment") +
  scale_y_continuous("Pr(decision)", breaks = seq(0, 1, by = 0.2)) +
  facet_grid2(
    desc ~ paste0("RD = ", rd) ,
    labeller = labeller(desc = label_wrap_gen(30))) + 
  theme_minimal() +
  theme(
    legend.position = "bottom",
    # this and the guide call below ensures that you get
    # legends for decision type and trt on different rows
    legend.box="vertical",
    legend.title = element_text(size = 8) ,
    strip.text.y.right = element_text(angle = 0,
                                      hjust = 0,
                                      vjust = 0.2,
                                      size = 7),
    axis.ticks = element_blank(),
    strip.text.x.top = element_text(size = 7),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey",
                                  linewidth = 0.1,
                                  linetype = 1),
    axis.title.y=element_text(size = 7),
    axis.text.y = element_text(size = 7), 
    axis.text.x = element_text(size = 7), 
    axis.title.x = element_text(size = 7),
    legend.text = element_text(size = 7)
  ) +
  guides(
    colour = guide_legend(nrow = 1),
    linetype = guide_legend(nrow = 1)
  )


```


{{< pagebreak >}}

## Sample size

@tbl-n-1 shows the expected number of participants on each treatment for a stopping decision (either NI or inferiority) to occur.

```{r, echo = F, eval = T}
#| label: sample_size_1
#| code-summary: Compute expected sample size


i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  # d_dec_1 <- copy(l[[i]]$d_pr_dec[par %in% c("rd_2_1", "rd_3_1")])
  # config for scenario
  l_cfg <- copy(l[[i]]$cfg)
  
  N_interim <- length(l_cfg$N_pt)
  N_trt <- length(l_cfg$btrt)
  # 
  d_grid <- CJ(
    sim = 1:l_cfg$nsim,
    ia = 1:N_interim,
    trt = 1:N_trt
  )
    
  d_all <- copy(l[[i]]$d_all)
  
  d_all <- merge(
    d_grid,
    d_all[, .(N = sum(N)), keyby = .(sim, ia, trt)],
    by = c("sim", "ia", "trt"),
    all.x = T
  )
  
  d_all[is.na(N), N := 0]
  d_all[, N_cmtv := cumsum(N), keyby = .(sim, trt)]
  # d_all[, N_cmtv := nafill(N_cmtv, type = "locf"), keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_all
    )
  )
  
}
```

```{r, eval = T}
#| echo: FALSE
#| label: tbl-n-1
#| tbl-cap: 'Expected number of participants by treatment group for each scenario'
#| tbl-pos: H

# DONT assume that all arms ran to the final interim. The result is just an
# artefact of imputation. Nevertheless it gives the true sample size by arm.
d_N_trt <- d_N_by_rand[, .SD[.N], keyby = .(scenario, desc, sim, trt)]
d_N_tot <- d_N_trt[, .(N_cmtv = sum(N_cmtv)), keyby = .(scenario, desc, sim) ]


d_tbl_1_cur <- rbind(
  d_N_trt[, .(N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc, trt)],
  d_N_tot[, .(trt = "all", N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc)]
)

d_tbl_1_cur[trt == 1, trt := "SoC"]
d_tbl_1_cur[trt == 2, trt := "Deferred"]
d_tbl_1_cur[trt == 3, trt := "Discontinued"]

# extract the true risk differences by comparison 
# we are only interested in Deferred vs soc and Discontinued vs soc
d_b_trt <- data.table()

for(i in 1:length(l)){
  l_cfg <- copy(l[[i]]$cfg)
  btrt <- unlist(l_cfg$btrt)
  d_tmp <- data.table(
    scenario = i, 
    trt = c("Deferred", "Discontinued"), 
    rd = c(btrt[2]-btrt[1], btrt[3]-btrt[1]))
  d_b_trt <- rbind(d_b_trt, d_tmp)
}

d_tbl_1_cur <- merge(
  d_tbl_1_cur, 
  d_b_trt, 
  by = c("scenario", "trt"),
  all.x = T)

d_tbl_1_cur <- dcast(
  d_tbl_1_cur, 
  scenario + desc ~ trt , 
  value.var = list("N_mu", "N_sd", "rd"))

d_tbl_1_cur[, `:=`(rd_SoC = NULL, rd_all = NULL)]

setcolorder(d_tbl_1_cur, c("scenario", "desc", "rd_Deferred", "rd_Discontinued"))

d_tbl_1_cur <- d_tbl_1_cur[order(scenario, desc)]
d_tbl_1_cur <- d_tbl_1_cur[, .SD, .SDcols = !c("scenario")]



g_tbl <- d_tbl_1_cur |> 
  gt(groupname_col = "desc")  |>
  fmt_number(
    columns = starts_with("N_"),
    decimals = 0, drop_trailing_zeros = TRUE) |>
  fmt_number(
    columns = starts_with("rd"),
    decimals = 2, drop_trailing_zeros = TRUE) |> 
  tab_spanner(
    label = md("Risk diff (relative to SoC)"),
    columns = starts_with("rd"),
  ) |>
  tab_spanner(
    label = md("Sample size, mean (sd)"),
    columns =  starts_with("N_mu")
  ) |>
  # cols_width(
  #   starts_with("N_mu") ~ pct(20),
  #   starts_with("rd") ~ pct(14)
  # ) |>
  cols_align(
    columns = starts_with("rd"),
    align = "center"
  ) |>
  cols_align(
    columns = starts_with("N_mu"),
    align = "center"
  )  |>
  cols_merge(
    columns = c("N_mu_Deferred", "N_sd_Deferred"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_Discontinued", "N_sd_Discontinued"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_SoC", "N_sd_SoC"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_all", "N_sd_all"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_label(
    rd_Deferred = "Deferred",
    rd_Discontinued = "Discontinued",
    N_mu_Deferred = "Deferred",
    N_mu_Discontinued = "Discontinued",
    N_mu_SoC = "SoC",
    N_mu_all = "Total"
  ) |>
  tab_options(
    # table.width = pct(100),
    table.font.size = pct(65)
  ) 


g_tbl
```


```{r, echo = F, eval = F}
#| label: sample_size_2
#| code-summary: Compute expected sample size for NI decision

# Expected number of pt to make a given decision type by arm for those arms
# for which the decision type is relevant.


i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_pr_dec[par %in% c("rd_2_1", "rd_3_1")])
  # config for scenario
  l_cfg <- copy(l[[i]]$cfg)
  
  N_interim <- length(l_cfg$N_pt)
  N_trt <- length(l_cfg$btrt)
  
  # The first decision by sim, comparison, irrespective of type. 
  # For example, if an NI happens first for rd_2_1 then that is it, irrespective
  # of what could have happened if enrolment had continued.
  
  d_dec_1 <- d_dec_1[dec == 1, .SD[1], keyby = .(sim, par)][
      , .(sim, par, ia, rule, dec)]
  
  
  # 
  d_grid <- CJ(
    sim = 1:l_cfg$nsim,
    ia = 1:N_interim,
    rule = c("inf", "ni"),
    par = c("rd_2_1", "rd_3_1")
  )
  d_grid[par == "rd_2_1", trt := 2]
  d_grid[par == "rd_3_1", trt := 3]
    
  
  d_dec_1 <- merge(
    d_grid ,
    d_dec_1, 
    by = c("sim", "par", "ia", "rule"),
    all.x = T
  )
  
  d_dec_inf <- merge(
    d_grid ,
    d_dec_inf, 
    by = c("sim", "par", "ia"),
    all.x = T
  )
  
  
  d_all <- copy(l[[i]]$d_all)
  
  d_all <- merge(
    d_grid,
    d_all[, .(N = sum(N)), keyby = .(sim, ia, trt)],
    by = c("sim", "ia", "trt"),
    all.x = T
  )
  
  d_all[is.na(N), N := 0]
  d_all[, N_cmtv := cumsum(N), keyby = .(sim, trt)]
  # d_all[, N_cmtv := nafill(N_cmtv, type = "locf"), keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_all
    )
  )
  
}
```



{{< pagebreak >}}
## Parameter estimation

@tbl-post-rd and @fig-expected-rd show the expected value of the posterior means (and 95% interval) for the treatment effects by scenario.

```{r, echo = F, eval = T}
#| label: post-means
#| code-summary: Distributions of posterior means (unconditional)

# Distribution of posterior means for parameters of interest.

# Some simulated trials will have stopped prior to the maximum sample size and
# these will have NA for their posterior means. If you were to summarise these 
# posterior means, they would be conditional on the trial having 'survived' 
# until the relevant interim. This means that you have missing data at later 
# interims, which creates a selection bias in that your selection of sims at any
# given interim are not a random sample, but rather a sample conditioned on the 
# stopping rules. 

# If you do not account for this in some way then a summary can be either 
# optimistic or pessimistic depending on how the stopping rules interact 
# with the data. Here we try to account for this missingness by imputing the 
# missing posterior means with locf within each simulation.
# Note that this is really only a partial 'fix' to get a sense of whether 
# our estimates is representative of the parameter values we used to simulate
# the data.

i <- 1
d_post_1 <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$cfg)
  
  b_trt <- unlist(l_cfg$btrt)
  
  # params
  d_pars <- copy(l[[i]]$d_post_smry_1)
  d_pars <- d_pars[par %in% c("rd_2_1", "rd_3_1"), .(sim, ia, par, mu)]
  d_pars[, mu :=  nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  # interim looks
  d_N <- data.table(ia = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  d_pars <- base::merge(d_pars, d_N, by = "ia")
  
  d_pars[par == "rd_2_1", rd_tru := b_trt[2] - b_trt[1]]
  d_pars[par == "rd_3_1", rd_tru := b_trt[3] - b_trt[1]]
  
  d_post_1 <- rbind(
    d_post_1,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_pars
      )
  )

}

d_post_1 <- d_post_1[order(scenario, desc, sim, ia)]

```

```{r, eval = T}
#| echo: FALSE
#| label: tbl-post-rd
#| tbl-cap: 'Parameter estimation - risk difference (expectation of posterior means and 95% interval)'
#| tbl-pos: H

d_tbl_1_cur <- d_post_1[,
                 .(mu = mean(mu),
                   q_025 = quantile(mu, prob = 0.025),
                   q_975 = quantile(mu, prob = 0.975)), 
                 keyby = .(scenario, desc, ia, par, rd_tru, N)]
# setorderv(d_fig, cols = "scenario", order = -1L)
d_tbl_1_cur[, desc := factor(desc, levels = unique(d_post_1$desc))]


d_tbl_1_cur <- dcast(d_tbl_1_cur, scenario + desc + par + rd_tru ~ N, value.var = list("mu", "q_025", "q_975"))

ci_names <- function(x = 500){
  paste0(c("mu_","q_025_", "q_975_"), x)
}
setcolorder(
  d_tbl_1_cur, 
  c("scenario", "desc",  "par", "rd_tru",
    ci_names(300), ci_names(400), ci_names(500), ci_names(600)))

d_tbl_1_cur[par == "rd_2_1", par := "Deferred"]
d_tbl_1_cur[par == "rd_3_1", par := "Discontinued"]
d_tbl_1_cur <- d_tbl_1_cur[, .SD, .SDcols = !c("scenario")]



g_tbl <- d_tbl_1_cur |>
  gt(groupname_col = "desc")  |>
  cols_align(
    columns = 1:2,
    align = "left"
  ) |>
  cols_align(
    columns = 3:ncol(d_tbl_1_cur),
    align = "right"
  )   |> 
  cols_merge(
    columns = c("par", "rd_tru"),
    pattern = "<<{1}>><< ({2})>>"
  )  |> 
  cols_merge(
    columns = c("mu_300", "q_025_300", "q_975_300"),
    pattern = "<<{1}>><< ({2}, {3})>>"
  )  |> 
  cols_merge(
    columns = c("mu_400", "q_025_400", "q_975_400"),
    pattern = "<<{1}>><< ({2}, {3})>>"
  )  |> 
  cols_merge(
    columns = c("mu_500", "q_025_500", "q_975_500"),
    pattern = "<<{1}>><< ({2}, {3})>>"
  )  |> 
  cols_merge(
    columns = c("mu_600", "q_025_600", "q_975_600"),
    pattern = "<<{1}>><< ({2}, {3})>>"
  )  |>
  cols_label(
    par = "Comparison (true RD)",
    mu_300 = "300",
    mu_400 = "400",
    mu_500 = "500",
    mu_600 = "600"
  )  |>
  tab_spanner(
    label = html("Risk difference (expectation of posterior means and 95 pct interval)"),
    columns = 2:ncol(d_tbl_1_cur)
    ) |>
  tab_options(
    table.font.size = pct(65),
    latex.use_longtable = TRUE,
    latex.header_repeat = TRUE 
    ) |>
   fmt_number(decimals = 2, drop_trailing_zeros = F)

g_tbl

```

```{r, eval = T, echo = F}
#| label: fig-expected-rd
#| fig-cap: 'Distribution of posterior means for risk difference treatment effects by interim and  simulation scenario'
#| fig-height: 8
#| fig-width: 7
#| fig-pos: H

d_fig <- copy(d_post_1)
d_fig[, desc := factor(desc, levels = unique(d_fig$desc))]
d_fig[, N := factor(N)]

d_fig[par == "rd_2_1", par := paste0("Deferred")]
d_fig[par == "rd_3_1", par := paste0("Discontinued")]


ggplot(data = d_fig, aes(x = N, y = mu, col = par)) +
  geom_boxplot(outlier.size = 0.1, lwd = 0.3) +
  ggh4x::facet_grid2(desc ~ . , 
             labeller = labeller(
               desc = label_wrap_gen(25)
               # ,
               # par = label_wrap_gen(15)
               ), 
             scales = "free",
             axes = "y",
             independent = "y")  +
  scale_color_manual("", values = c("#E66100", "#5D3A9B")) +
  scale_x_discrete("") +
  scale_y_continuous("Risk difference") +
  theme_bw() +
  theme(
    legend.position = "bottom",
        text = element_text(size = 6),
        strip.text.y.right = element_text(angle = 0,
                                      hjust = 0,
                                      vjust = 0.2,
                                      size = 6),
        strip.text.x = element_text(angle = 0, size = 6),
        axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size = 6),
        axis.text.y = element_text(size = 6),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey",
                                  linewidth = 0.1,
                                  linetype = 1))


```

## Observed proportion with treatment success

@tbl-obs-prop shows the observed proportion with treatment success by scenario, strata and treatment arm.

```{r, echo = F, eval = T}
#| label: empirical-risk
#| code-summary: Summaries of empirical probability of treatment success

i <- 1
d_prop_success <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$cfg)
  
  # interim looks
  d_enrolment <- data.table(ia = seq_along(l_cfg$N_pt), N_enrol = cumsum(l_cfg$N_pt))
  
  # enrolment
  d_N <- data.table(ia = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))

  # observed data
  d_all <- copy(l[[i]]$d_all)
  d_all <- base::merge(d_all, d_enrolment , by = "ia")
  d_all <- d_all[, .(y = sum(y), N = sum(N)), keyby = .(trt, lung, pseudo)]
  d_all[, p_obs := y / N]
  
  d_all <- dcast(d_all, lung + pseudo ~ trt, value.var = "p_obs")
  
  d_all[, rd_2_1 := `2` - `1`]
  d_all[, rd_3_1 := `3` - `1`]
  
  d_prop_success <- rbind(
    d_prop_success, 
    cbind(
      scenario = i, 
      desc = l_cfg$desc, 
      d_all
      )
  )
  
}

setnames(d_prop_success, paste0(1:3), paste0("p_obs_", 1:3))

```


```{r}
#| echo: FALSE
#| label: tbl-obs-prop
#| tbl-cap: 'Observed proportion of failures and differences by scenario, treatment and strata'
#| tbl-pos: H

d_tbl_1_cur <- copy(d_prop_success)
# setorderv(d_fig, cols = "scenario", order = -1L)
d_tbl_1_cur[, desc := factor(desc, levels = unique(d_post_1$desc))]

d_tbl_1_cur[, lung := factor(lung, levels = 1:2, labels = c("< 70% pFEV1", ">= 70% pFEV1"))]
d_tbl_1_cur[, pseudo := factor(pseudo, levels = 1:2, labels = c("no pseudo", "pseudo"))]

d_tbl_1_cur[, `:=`(scenario = NULL)]





g_tbl <- d_tbl_1_cur |>
  gt(groupname_col = "desc")  |>
  cols_align(
    columns = 1:3,
    align = "left"
  ) |>
  cols_align(
    columns = 4:ncol(d_tbl_1_cur),
    align = "right"
  )   |> 
  cols_label(
    lung = "Lung status",
    pseudo = "Colonisation",
    p_obs_1 = "SoC",
    p_obs_2 = "Deferred",
    p_obs_3 = "Discontinued",
    rd_2_1 = "Deferred",
    rd_3_1 = "Discontinued"
  )  |>
  tab_spanner(
    label = html("Proportion with treatment failure"),
    columns = starts_with("p_obs")
    ) |>
  tab_spanner(
    label = html("Difference in proportions"),
    columns = starts_with("rd_")
    ) |>
  tab_options(
    table.font.size = pct(65),
    latex.use_longtable = TRUE,
    latex.header_repeat = TRUE 
    ) |>
   fmt_number(decimals = 2, drop_trailing_zeros = F)

g_tbl

```


{{< pagebreak >}}
# Repository status {.unlisted .unnumbered}

\footnotesize
```{r}
#| echo: false
repo <- repository(path = ".")
summary(repo)
```
\normalsize


{{< pagebreak >}}
## References

<!-- Needs to have a citation for this to work otherwise you will get the old \end{CSLReferences} error  -->


