---
title: BEAT CF MARS
subtitle: |
    Simulation Report
date: last-modified
date-format: "D MMMM YYYY"
author: Mark Jones
editor: source
bibliography: ../etc/refs.bib
csl: ../etc/elsevier-harvard.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r}
#| label: load-data
#| code-summary: Data load
#| echo: false

# uml digs
suppressPackageStartupMessages(library(nomnoml))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(qs))
suppressPackageStartupMessages(library(git2r))
suppressPackageStartupMessages(suppressWarnings(library(gt)))
suppressPackageStartupMessages(library(ggh4x))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(pbapply))


toks <- unlist(tstrsplit(getwd(), "/")) 
if(toks[length(toks)] == "beatcf-mars-sim"){
  prefix_cfg <- "./etc/sim15/"
  prefix_stan <- "./stan"
  prefix_fig <- "./fig"
  prefix_data <- "./data"
  prefix_r <- "./R"
} else {
  prefix_cfg <- "../etc/sim15/"
  prefix_stan <- "../stan"
  prefix_fig <- "../fig"
  prefix_data <- "../data"
  prefix_r <- "../R"
}



source(paste0(prefix_r, "/libs.R"))
source(paste0(prefix_r, "/init.R"))
source(paste0(prefix_r, "/util.R"))
source(paste0(prefix_r, "/data-sim15.R"))


# Each input file corresponds to the results from a single simulation
# scenario/configuration.
# Load all the files into a single list.

# files of interest
sim_lab <- "sim15-02"

flist <- list.files(paste0(prefix_data, "/", sim_lab), pattern = "sim15")
toks <- list()
l <- list()
i <- 1

for(i in 1:length(flist)){

  l[[i]] <- qs::qread(file.path(paste0(prefix_data, "/", sim_lab), flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

N_sims <- l[[1]]$cfg$nsim



# temp to replace
# l_spec <- get_prototype_cfg()
l_spec_1 <- l[[1]]$l_spec

desc_labs <- character()

for(i in 1:length(l)){
  
  desc_labs <- c(desc_labs, l[[i]]$l_spec$desc)
  
}

```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  BEAT-CF MARS |
|Reference intervention: |  Standard airway clearance therapy at least twice per day for 14 days, plus 14 days of oral antibiotics commencing as soon as practicable (and ≥ 72 hours and < 7 days) after onset of new or acute worsening of cough. |   
|Intervention: | (1) Deferred antibiotic strategy (2) Early discontinuation antibiotic strategy. |   
|Outcome: | Time to recovery from onset of exacerbation |   
|Study design:  |   Bayesian adaptive trial with early stopping rules | 
|Sponsor:  |    University of Sydney, NSW, Australia | 
Protocol: |  todo |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Tom Snelling | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}

| Version    |   Date     | Change    |   Reason     |
|:----|:------------|:----|:------------|
| 0.1 | 2026-03-04 | First version revised design | N/A |
| 0.2 | 2026-03-04 | Incorporate carry-over effects into the healthy state | N/A |


{{< pagebreak >}}


# Design

State transition 
Joint model

# Model

Narrative overview

# Data Structure

What data we need


# Decision Processes

Reference the config

# Results

## Example Trial

The example is from the first scenario and takes the first simulation.
All the figures in this section relate to a single trial.

@fig-disc-eh-ex-1 shows the distribution of the number of exacerbations each participant has over the `r l_spec_1$followup` day followup.

@fig-post-par-ex-1 shows the prior and posterior distribution on the parameters from the model at the time of the last analysis.
These include the intercept terms for the piecewise baseline hazard, the covariate effects of baseline ppfev1, treatment effects and variance components used to account for repeat measures.

@fig-post-soj-ex-1 shows the prior and posterior on the sojourn and difference in sojourn times in each state (**H**ealthy and **E**xacerbation).
To derive these, we use the model joint posterior to simulate the trajectory for a set of participants. 
Key points include:

1. We simulate a trial population using the same number as presently observed in the analysis for each draw from the model parameters joint posterior
2. Everyone starts in an exacerbation state and their state is updated by day
3. The decision horizon is set to `r l_spec_1$decision_fu` days
4. We assume a fixed treatment policy over the decision horizon and the baseline covariates are resampled from the actual trial population
5. For each participant in the population, we compute the total days in an exacerbation state
6. For each simulated population we compute the mean days in exacerbation state

Point (4) makes the strong assumption that while carry over effects of a policy may exist, their is no cumulative impact beyond that.
Point (6) gives us a distribution of mean days in exacerbation state with the uncertainty propagated from the posterior.

The equivalent process can be run to obtain a view on the prior for these parameters.

@tbl-soj-ex-01 summarises the distribution of mean days in exacerbation and health state under each treatment arm for this specific example.
@tbl-diff-ex-01 summarises the difference between each treatment arm and soc.

```{r}
#| label: fig-disc-eh-ex-1
#| fig-cap: 'Distribution of exacerbations'
#| fig-height: 4
#| fig-width: 4
#| out-height: 4in
#| out-width: 4in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

# Distribution of exacerbations in this sample.

l_cfg <- copy(l[[1]]$l_spec)
d_w <- copy(l[[1]]$d_w[sim == l_cfg$ex_trial_ix[1]])

# number of unique elements in vec
d_fig <- d_w[, .(
  n_E = uniqueN(per_id[state == "E"])
  ), keyby = .(sim, id)
]

# d_fig[, .N, keyby = .(desc, sim, n_E)]

# https://github.com/tidyverse/ggplot2/issues/2051
ggplot(d_fig, aes(x = n_E)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_x_continuous("Exacerbations", breaks = 0:10) +
  scale_y_continuous("Proportion of Pts", 
                     breaks = seq(0, 1, by = 0.1))
```


```{r}
#| label: fig-enrol-ex-1
#| fig-cap: 'Allocation to each treatment arm'
#| fig-height: 4
#| fig-width: 4
#| out-height: 4in
#| out-width: 4in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

# Distribution of exacerbations in this sample.

l_cfg <- copy(l[[1]]$l_spec)
d_w <- copy(l[[1]]$d_w[sim == l_cfg$ex_trial_ix[1]])

d_w[, .N, keyby = .(tstart, trt)]

# number of unique elements in vec
d_fig <- d_w[, .(
  n_E = uniqueN(per_id[state == "E"])
  ), keyby = .(sim, id)
]

# d_fig[, .N, keyby = .(desc, sim, n_E)]

# https://github.com/tidyverse/ggplot2/issues/2051
ggplot(d_fig, aes(x = n_E)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_x_continuous("Exacerbations", breaks = 0:10) +
  scale_y_continuous("Proportion of Pts", 
                     breaks = seq(0, 1, by = 0.1))
```

```{r}
#| label: fig-post-par-ex-1
#| fig-cap: 'Prior and posterior for model parameters'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)


d_pri_par <- copy(l[[1]]$d_pri_par[sim == l_cfg$ex_trial_ix[1]])
d_post_par <- copy(l[[1]]$d_post_par[sim == l_cfg$ex_trial_ix[1]])

# only get the last one:

d_pri_par <- d_pri_par[ic == max(ic)]
d_post_par <- d_post_par[ic == max(ic)]

d_pri_par[, `:=`(ic = NULL)]
d_post_par[, `:=`(ic = NULL)]

d_pri_par[, desc := "prior"]
d_post_par[, desc := "posterior"]

d_par <- rbind(d_pri_par, d_post_par)

d_fig <- melt(d_par, id.vars = c("desc", "sim"))
d_fig[, desc := factor(desc, levels = c("posterior", "prior"))]

p_1 <- ggplot(d_fig[variable %like% "a_"], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 4)

p_2 <- ggplot(d_fig[!(variable %like% "a_")], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 4)

p_1 / p_2 +
  plot_layout(heights = c(1, 1), guides = "collect")

```


```{r}
#| label: fig-post-soj-ex-1
#| fig-cap: 'Prior and posterior for sojourn times'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)

d_pri_res <- copy(l[[1]]$d_pri_res[sim == l_cfg$ex_trial_ix[1]])
d_post_res <- copy(l[[1]]$d_post_res[sim == l_cfg$ex_trial_ix[1]])

# only get the last one:

d_pri_res <- d_pri_res[ic == max(ic)]
d_post_res <- d_post_res[ic == max(ic)]

d_pri_res[, `:=`(ic = NULL)]
d_post_res[, `:=`(ic = NULL)]

d_pri_res[, desc := "prior"]
d_post_res[, desc := "posterior"]

d_par <- rbind(d_pri_res, d_post_res)

d_fig <- melt(d_par, id.vars = c("desc", "sim"))
d_fig[, desc := factor(desc, levels = c("posterior", "prior"))]

p_1 <- ggplot(d_fig[variable %like% "mu_"], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 3)

p_2 <- ggplot(d_fig[!(variable %like% "mu_")], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 2)

p_1 / p_2 +
  plot_layout(heights = c(1, 1), guides = "collect")
```


```{r, echo = F}

l_cfg <- copy(l[[1]]$l_spec)
d_trt <- copy(l[[1]]$d_trt_effects[sim == l_cfg$ex_trial_ix[1]])
d_pr_dec <- copy(l[[1]]$d_pr_dec[sim == l_cfg$ex_trial_ix[1]])

d_pr_dec <- dcast(d_pr_dec, sim + ic + trt ~ rule, value.var = c("p"))
setnames(d_pr_dec, "fut", "pr_fut")
setnames(d_pr_dec, "ni", "pr_ni")

d_trt[, N := cumsum(l_cfg$N_pt)[ic]]
d_trt[, desc := l_cfg$desc]

d_trt[par %like% "_E", state := "Exacerbation"]
d_trt[par %like% "_H", state := "Healthy"]
d_trt[par %like% "_def_", trt := "def"]
d_trt[par %like% "_dis_", trt := "dis"]
d_trt[par %like% "_soc_", trt := "soc"]

d_trt[par %like% "del_",  stat := "diff"]
d_trt[par %like% "mu_",  stat := "mean"]

d_trt[state == "Exacerbation", out := sprintf("%.1f (%.1f, %.1f)", mu, lwr, upr)]
d_trt[state == "Healthy", out := sprintf("%.1f (%.1f, %.1f)", mu, lwr, upr)]


d_trt <- base::merge(d_trt, d_pr_dec, by = c("sim", "ic", "trt"), all.x = T)

```

```{r, echo = F, eval = T}
#| label: tbl-soj-ex-01
#| tbl-cap: 'Posterior mean and 95% CrI sojourn times'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

d_tbl <- dcast(d_trt[stat == "mean"], desc + state + trt ~ N, value.var = "out")

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 9) 
```


```{r, echo = F, eval = T}
#| label: tbl-diff-ex-01
#| tbl-cap: 'Posterior difference in sojourn times'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

d_tbl <- d_trt[stat == "diff"]
d_tbl <- d_tbl[order(trt, N)]

kableExtra::kbl(
  d_tbl[state == "Exacerbation", .(desc, N, trt, out, pr_fut, pr_ni)], 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "\\addlinespace"),
  col.names=c("Scenario", "N", "Treatment", "Differnce", "Pr(fut)", "Pr(NI)"), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 9) 
```


## Trial Population

```{r, eval = T}
#| label: exac-comp
#| code-summary: Frequency of exacerbation
#| echo: FALSE

i <- 1
d_w <- data.table()

g_scen_names <- c()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # add desc...
  d_tmp <- copy(l[[i]]$d_w)
  d_tmp[, desc := l_cfg$desc]
  
  g_scen_names <- c(g_scen_names, l_cfg$desc)
  
  d_tmp[, ix_desc := i]
  d_w <- rbind(d_w, d_tmp)
}

d_w[, desc := factor(desc, levels = g_scen_names)]
d_w[, len_seg := tstop - tstart]


# Pick out trial data from one simulation for each scenario
ix_smpl <- 1:5 # sort(sample(1:max(d_w$sim), 5))
d_w_smpl <- d_w[sim %in% ix_smpl, ]
```

@fig-dist-n-eh shows the distribution of number of exacerbations for a random selection of simulations and @fig-dist-dur-eh shows the distribution of exacerbation days over all exacerbations that occurred.

@fig-dist-dur_he shows the distribution of health days for the first healthy period (all participants are assumed to enter in a healthy state and may progress to one or more exacerbations).
@fig-dist-n-trt shows the distribution of treatment allocations over the entire follow up period.
@fig-dist-n-enrol shows the distribution of enrolment over the all simulations for each scenario.

```{r}
#| label: fig-dist-n-eh
#| fig-cap: 'Number of exacerbations in trial population over 365 days followup for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[, .(
  n_E = uniqueN(per_id[state == "E"])
  ), keyby = .(desc, sim, id)
]

# d_fig[, .N, keyby = .(desc, sim, n_E)]

# https://github.com/tidyverse/ggplot2/issues/2051
ggplot(d_fig, aes(x = n_E)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_x_continuous("Exacerbations", breaks = 0:10) +
  scale_y_continuous("Proportion of Pts", 
                     breaks = seq(0, 1, by = 0.1)) +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl))

```


```{r}
#| label: fig-dist-dur-eh
#| fig-cap: 'Duration of exacerbations in trial population over 365 days followup for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "E", .(dur_eh = sum(len_seg)), keyby = .(desc, sim, id, per_id)]

ggplot(d_fig, aes(x = dur_eh)) +
  geom_histogram(bins = 20) +
  scale_x_continuous("Duration (days)") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl))

```


```{r}
#| label: fig-dist-dur_he
#| fig-cap: 'Duration of first healthy period in trial population for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "H" & per_id == 1, .(dur_he_1 = sum(len_seg)), keyby = .(desc, sim, id)]

ggplot(d_fig, aes(x = dur_he_1)) +
  geom_histogram(bins = 20) +
  scale_x_continuous("Duration (days)") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl))

```


```{r}
#| label: fig-dist-n-trt
#| fig-cap: 'Distribution of treatment allocation (randomisation of each treatment arm over trial population)'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "E", .N, keyby = .(desc, sim, id, per_id, trt)]
d_fig[, `:=`(N = NULL)]
d_fig <- d_fig[, .N, keyby = .(desc, sim, trt)]
d_fig[, trt := factor(trt, levels = l_spec_1$trt_lab)]
# d_fig[, sum(N), by = sim]

ggplot(d_fig, aes(x = trt, y = N)) +
  geom_bar(stat = "identity") +
  scale_x_discrete("") +
  scale_y_continuous("N") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl))

```



```{r}
#| label: fig-dist-n-enrol
#| fig-cap: 'Distribution of enrolment size over all simulations'
#| fig-height: 3
#| fig-width: 6
#| out-height: 3in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

# unique pt enrolled by sim
d_fig <- d_w[, .(N = length(unique(id))), keyby = .(desc, sim)]
d_fig <- d_fig[, .(freq = .N), keyby = .(desc, N)]

# d_fig[, sum(N), by = sim]

ggplot(d_fig, aes(x = N, y = freq)) +
  geom_bar(stat = "identity", width = 10) +
  scale_x_continuous("Enrolment", breaks = cumsum(l_spec_1$N_pt)) +
  scale_y_continuous("N") +
  facet_wrap(desc ~ ., nrow = 1)
```

## Parameters 

Expectation and uncertainty of posterior means for model parameter estimates.

```{r, eval = T}
#| label: tbl-par-comp
#| tbl-cap: 'Posterior parameter summary'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

i <- 1
d_par <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_tmp <- copy(l[[i]]$d_post_smry_1[
    par %in% l_spec_1$par_names])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_tmp[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  d_smry <- d_tmp[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_par <- rbind(
    d_par,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_par[, stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]

d_tbl <- dcast(d_par, desc + par ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl <- d_tbl[order(desc, par)]

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("llccc"),
  linesep = "",
  col.names=c("Scenario", "Parameter", cumsum(l_cfg$N_pt)), 
  digits = 1)  |>
  kableExtra::kable_styling(font_size = 8) |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3:5, width = "3.1cm") 

```

## Sojourn Time

@tbl-soj-01 shows the expectations and 95% uncertainty interval for the posterior means of the sojourn times (days) in each health state by treatment arm over a simulated `r l_spec_1$decision_fu` day decision horizon.
That is, we collect the posterior means from each simulation and analysis and summarise them by the mean and the 2.5 and 97.5 percentiles.

@tbl-soj-02 provides the difference in sojourn times by health state in each treatment arm relative to the control arm.

```{r, echo = F, eval = T}
i <- 1
d_soj <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_par <- copy(
    l[[i]]$d_trt_effects[par %in% c(
      "mu_soc_H", "mu_def_H", "mu_dis_H",
      "mu_soc_E", "mu_def_E", "mu_dis_E"
      )])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_par[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  # only summarising the means
  d_smry <- d_par[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_soj <- rbind(
    d_soj,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_soj[par %like% "_E", state := "Exacerbation"]
d_soj[par %like% "_H", state := "Healthy"]
d_soj[state == "Exacerbation", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]
d_soj[state == "Healthy", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]

d_soj[par %like% "_def_", trt := "def"]
d_soj[par %like% "_dis_", trt := "dis"]
d_soj[par %like% "_soc_", trt := "soc"]

d_tbl <- dcast(d_soj, desc + state + trt ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, trt := factor(trt, levels = l_spec_1$trt_lab)]
d_tbl[, state := factor(state, levels = c("Healthy", "Exacerbation"))]
d_tbl <- d_tbl[order(desc, trt, state)]
```


```{r, echo = F, eval = T}
#| label: tbl-soj-01
#| tbl-cap: 'Expectation of posterior mean sojourn times by scenario, health state and treatment arm'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "",  "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 1)  |>
  kableExtra::kable_styling(font_size = 9) 
```

```{r, echo = F, eval = T}
i <- 1
d_soj <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_par <- copy(
    l[[i]]$d_trt_effects[par %in% c(
      "del_def_H", "del_dis_H", 
      "del_def_E", "del_dis_E"
      )])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_par[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  # only summarising the means
  d_smry <- d_par[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_soj <- rbind(
    d_soj,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_soj[par %like% "_E", state := "Exacerbation"]
d_soj[par %like% "_H", state := "Healthy"]
d_soj[state == "Exacerbation", stat := sprintf("%.2f (%.2f, %.2f)", mu, lo, hi)]
d_soj[state == "Healthy", stat := sprintf("%.2f (%.2f, %.2f)", mu, lo, hi)]

d_soj[par %like% "_def_", trt := "def"]
d_soj[par %like% "_dis_", trt := "dis"]
d_soj[par %like% "_soc_", trt := "soc"]

d_tbl <- dcast(d_soj, desc + state + trt ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, trt := factor(trt, levels = l_spec_1$trt_lab)]
d_tbl[, state := factor(state, levels = c("Healthy", "Exacerbation"))]
d_tbl <- d_tbl[order(desc, trt, state)]
```


```{r, echo = F, eval = T}
#| label: tbl-soj-02
#| tbl-cap: 'Expectation of difference in mean sojourn times by scenario, health state and treatment arm relative to the control arm'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 1)  |>
  kableExtra::kable_styling(font_size = 9) 
```



{{< pagebreak >}}


## Decision Probability

@tbl-dec-pr-1 shows the cumulative probability of the non-inferiority and futility decisions based on the difference in the exacerbation sojourn times.

```{r, echo = F, eval = T}
#| label: tbl-dec-pr-1
#| tbl-cap: 'Cumulative probability of each decision type'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H


i <- 1
d_dec <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_pr_dec[trt %in% c("def", "dis")])
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  
  # number of enrolments at each interim (interim sample size sequence)
  d_N <- data.table(ic = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  d_dec_1[, cdec := as.integer(cumsum(dec)>0), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := nafill(cdec, type = "locf"), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := as.logical(cdec)]
  
  d_dec_1 <- base::merge(d_dec_1, d_N, by = "ic")
  
  # cumulative proportion for which each decision quantity has been met by 
  # analysis and domain
  d_dec_1 <- d_dec_1[, .(pr_val = mean(cdec)), keyby = .(ic, N, rule, trt)]
  
  d_dec <- rbind(
    d_dec,
    cbind(scenario = i, desc = l_cfg$desc, d_dec_1)
  )

}

d_tbl <- dcast(
  d_dec, desc + rule  + trt ~ N, value.var = "pr_val")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, rule := factor(rule, levels = c("ni", "fut"), labels = c("NI", "Fut"))]
d_tbl <- d_tbl[order(desc, rule)]


kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "\\addlinespace"),
  col.names=c("Scenario", "Rule", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 11) |>
  kableExtra::column_spec(1, width = "4cm")|>
  kableExtra::column_spec(2:6, width = "1.5cm")



```





{{< pagebreak >}}

## Sample Size

Expected number of participant randomisations on each treatment arm under each scenario.

```{r, echo = F, eval = F}
#| label: sample-size-comp
#| code-summary: Compute expected sample size


i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  # l_cfg$desc
  
  N_analysis <- length(l_cfg$N_pt)
  N_trt <- 3
  # 
  d_grid <- CJ(
    sim = 1:l_cfg$nsim,
    ic = 1:N_analysis,
    trt = 1:N_trt
  )
  d_grid$trt <- l[[i]]$l_spec$trt_lab[d_grid$trt]
    
  d_all <- copy(l[[i]]$d_all[state == "E"])
  
  # time of last enrolment entering into interim analysis by sim
  d_analy_set <- copy(l[[i]]$d_analys_sets)
  d_analy_set <- d_analy_set[!is.na(t0_last), ]
  d_analy_set[, ic := 1]
  d_analy_set[, ic := cumsum(ic), by = sim]
  
  setkey(d_all, sim, t0)
  setkey(d_analy_set,  sim, t0_last)

  
  # identify the analysis that each pt first enters
  # Rolling join: find first cutoff_date >= enrol_date
  d_all[
    d_analy_set,
    on = .(sim, t0 = t0_last),
    roll = -Inf,
    ic := i.ic
  ]
  # last one carried back
  d_all[, ic := nafill(ic, type = "nocb"), by = sim]
  
  # how many randomisations per interim and trt arm (pt may have more than 
  # one randomisation over their followup)
  d_smry <- base::merge(
    d_grid,
    d_all[, .(.N), keyby = .(sim, ic, trt)],
    by = c("sim", "ic", "trt"),
    all.x = T
  )
  
  # some arms may stop after interim 1 and so there would be no count for 
  # subsequent interims
  d_smry[is.na(N), N := 0]
  d_smry[, N_cmtv := cumsum(N), keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_smry
    )
  )
  
}

d_N_trt <- d_N_by_rand[, .SD[.N], keyby = .(scenario, desc, sim, trt)]
d_N_tot <- d_N_trt[, .(N_cmtv = sum(N_cmtv)), keyby = .(scenario, desc, sim) ]

# add in total sample size
d_tbl <- rbind(
  d_N_trt[, .(N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc, trt)],
  d_N_tot[, .(trt = "all", N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc)]
)
setkey(d_tbl, scenario, desc)


kableExtra::kbl(
  d_tbl[, .SD, .SDcols = -c("scenario")], 
  col.names = c("Desc", "TRT", "Mean", "SD"),
  format = "simple", digits = 1)

```



{{< pagebreak >}}
# References {.unnumbered}

<!-- Needs to have a citation for this to work otherwise you will get the old \end{CSLReferences} error  -->

::: {#refs}
:::

{{< pagebreak >}}

# Appendix - Frailty Model {.unnumbered .appendix}

To speed up the model fit for the frailty model the frailty is marginalised as discussed in [@Ibrahim2001, Section 4.1.1].

Specifically, first assume participants $i$ with observations $j = 1, \dots, n_i$ and a frailty $u_i \sim \text{Gamma}(a, a)$ with mean 1 and variance $1/a$. The **conditional** likelihood for all observations from participant $i$ is:


$$
\begin{aligned}
\mathcal{L}_i(u_i) = \prod_{j=1}^{n_i} f(y_{ij} | u_i)
\end{aligned}
$$

with Weibull (PH) pdf $f$ and conditional hazard 

$$
\begin{aligned}
h(y_i \mid u_i) = 
\underbrace{u_i}_{\text{frailty}} \cdot 
\underbrace{\gamma}_{\text{baseline scale}} \cdot 
\underbrace{\alpha y_i^{\alpha-1}}_{\text{shape}} \cdot 
\underbrace{\theta_{ij}}_{\text{covariate effects}}
\end{aligned}
$$

with scale modifying covariate terms for ppfev at baseline and treatment effects entering in through $\theta_{ij} = \exp(x_{ij}^\prime \beta)$.
Alternatively, just suck the baseline scale into the linear predictor:

$$
\begin{aligned}
h(y_i \mid u_i) = 
u_i \cdot 
\alpha y_i^{\alpha-1}
\underbrace{\mu_{ij}}_{\text{covariate effects incl. baseline scale}}
\end{aligned}
$$

where $\mu = \exp(b_0 + x_{ij}^\prime \beta)$.

Assuming we observe full event series for all participants, subject $i$ contributes:

$$
\begin{aligned}
\mathcal{L}_i(u_i) = \prod_{j=1}^{n_i} \alpha \cdot  u_i \mu_{ij} \cdot y_{ij}^{\alpha - 1} \cdot  \exp(-u_i \mu_{ij} y_{ij}^\alpha)
\end{aligned}
$$

collect up terms in $u_i$

$$
\begin{aligned}
\mathcal{L}_i(u_i) = u_i^{n_i} \cdot C_i \cdot \exp(-u_i S_i)
\end{aligned}
$$

where $C_i = \prod_{j=1}^{n_i} \alpha \mu_{ij} y_{ij}^{\alpha - 1}$ and $S_i = \sum_{j=1}^{n_i} \mu_{ij} y_{ij}^\alpha$.

We want the marginal likelihood, i.e. 

$$
\begin{aligned}
\mathcal{L}_i = \int_0^\infty \mathcal{L}_i(u_i) \cdot p(u_i) du_i
\end{aligned}
$$

where $p(u) = \frac{a^a}{\Gamma(a)} u^{a-1} \exp(-a u)$ is the prior density for $u_i$, here a Gamma distribution (shape and rate parameterisation assuming equal shape and rate):

$$
\begin{aligned}
\mathcal{L}_i = \int_0^\infty u_i^{n_i} \cdot C_i \cdot \exp(-u_i S_i) \cdot \frac{a^a}{\Gamma(a)} u_i^{a-1} \exp(-a u_i) du_i
\end{aligned}
$$


Grouping all the $u_i$ terms together

$$
\begin{aligned}
\mathcal{L}_i = C_i \cdot \frac{a^a}{\Gamma(a)} \int_0^\infty u_i^{(n_i + a)-1} \cdot \exp(-(S_i + a)u_i) du_i
\end{aligned}
$$

Now we just do the old stats trick of recognising a stripped down $\text{Gamma}(c, r)$ pdf (missing the $\frac{c^r}{\Gamma{c}}$ terms) with the form $u^{c-1} \exp(-r u)$, so we know that the integrand of this must equal $\frac{\Gamma{c}}{c^r}$ in order for the density to integrate to one (see @Wackerly2007 for examples).

If we substitute with $c = n_i + 1$ and $r = S_i + a$, we get:

$$
\begin{aligned}
\mathcal{L}_i = C_i \cdot \frac{a^a}{\Gamma(a)} \frac{\Gamma(n_i + a)}{(S_i + a)^{(n_i + a)}}
\end{aligned}
$$

Taking logs gives what we need to incorporate into the stan model:

$$
\begin{aligned}
\log(\mathcal{L}_i) = \log(C_i) + a\log(a) - \log(\Gamma(a)) + log(\Gamma(n_i + a) - (n_i + a) \log(S_i + a)
\end{aligned}
$$

where the $C_i = \sum_j \log(\alpha \mu_{ij} y_{ij}^{\alpha-1}) = \sum_j (\log(\alpha ) + \log(\mu_{ij}) + (\alpha-1) \log(y_{ij}))$

What about if there are censored observations?
In this case the conditional likelihood is 

$$
\begin{aligned}
\mathcal{L}_i(u_i) = \prod_{j=1}^{n_i} f(y_{ij} | u_i) \cdot S(c_i | u_i)
\end{aligned}
$$

where the density is for the observed records and the survival function for the final sojourn time of duration $c_i$.
This works through as previously described and the marginal log-likelihood comes out as:

$$
\begin{aligned}
\log(\mathcal{L}_i) &= \sum_{j=1}^{n_i} [\log(\alpha) + \log(\mu_{ij}) +  (\alpha-1) \log(y_{ij}) ] + \\ 
 & \quad \log(\Gamma(n_i + a)) - \\ 
 & \quad \log(\Gamma(a)) + a \log(a) - \\
 & \quad (n_i + a)\log(S_i + a)
\end{aligned}
$$

where $S_i = \sum_{j = 1}^{n_i} \mu_{ij} y_{ij}^\alpha + \mu_{i,last} c_i^\alpha$ and where $n_i$ is now counting only the completed transitions, not the censored final sojourn.

In the above you can think of the first part as the log of thew weibull hazard at each time, what you would have if there was no frailty. 
The last part gives the updated rate parameter after pooling the prior rate with the cumulative hazard.




