---
title: BEAT CF MARS
subtitle: |
    Simulation Report
date: last-modified
date-format: "D MMMM YYYY"
author: Mark Jones
editor: source
bibliography: ../etc/refs.bib
csl: ../etc/elsevier-harvard.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r}
#| label: load-data
#| code-summary: Data load
#| echo: false

# uml digs
suppressPackageStartupMessages(library(nomnoml))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(qs))
suppressPackageStartupMessages(library(git2r))
suppressPackageStartupMessages(suppressWarnings(library(gt)))
suppressPackageStartupMessages(library(ggh4x))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(pbapply))


toks <- unlist(tstrsplit(getwd(), "/")) 
if(toks[length(toks)] == "beatcf-mars-sim"){
  prefix_cfg <- "./etc/sim15/"
  prefix_stan <- "./stan"
  prefix_fig <- "./fig"
  prefix_data <- "./data"
  prefix_r <- "./R"
} else {
  prefix_cfg <- "../etc/sim15/"
  prefix_stan <- "../stan"
  prefix_fig <- "../fig"
  prefix_data <- "../data"
  prefix_r <- "../R"
}



source(paste0(prefix_r, "/libs.R"))
source(paste0(prefix_r, "/init.R"))
source(paste0(prefix_r, "/util.R"))
source(paste0(prefix_r, "/data-sim15.R"))


# Each input file corresponds to the results from a single simulation
# scenario/configuration.
# Load all the files into a single list.

# files of interest
sim_lab <- "sim15-02"

flist <- list.files(paste0(prefix_data, "/", sim_lab), pattern = "sim15")
toks <- list()
l <- list()
i <- 1

for(i in 1:length(flist)){

  l[[i]] <- qs::qread(file.path(paste0(prefix_data, "/", sim_lab), flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

N_sims <- l[[1]]$cfg$nsim



# temp to replace
# l_spec <- get_prototype_cfg()
l_spec_1 <- l[[1]]$l_spec

desc_labs <- character()

for(i in 1:length(l)){
  
  desc_labs <- c(desc_labs, l[[i]]$l_spec$desc)
  
}

```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  BEAT-CF MARS |
|Reference intervention: |  Standard airway clearance therapy at least twice per day for 14 days, plus 14 days of oral antibiotics commencing as soon as practicable (and ≥ 72 hours and < 7 days) after onset of new or acute worsening of cough. |   
|Intervention: | (1) Deferred antibiotic strategy (2) Early discontinuation antibiotic strategy. |   
|Outcome: | Time to recovery from onset of exacerbation |   
|Study design:  |   Bayesian adaptive trial with early stopping rules | 
|Sponsor:  |    University of Sydney, NSW, Australia | 
Protocol: |  todo |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Tom Snelling | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}


\tiny
| Version    |   Date     | Change    |   Reason     |
|:----|:----|:--------|:--------|
| 0.1 | 2026-03-04 | First version revised design | - |
| 0.2 | 2026-03-04 | Incorporate carry-over effects into the healthy state | Feedback from CI |
| 0.2 | 2026-05-18 | Discrete time process | Better representation of state |
\normalsize

{{< pagebreak >}}

# Design

Beat CF Mars is a three arm, sequential trial with adaptive stopping rules based on non-inferiority and futility.
The primary endpoint is the expected exacerbation days within a 90 day window.

Participants are enrolled in a well state and followed up over 365 days.
On exerperiencing the first exacerbation in the follow up period, participants are randomised to:

+ standard of care antibiotics (soc)
+ deferred antibiotics (def)
+ early discontinue antibiotics (dis)

and tracked for resolution of symptoms.
In subsequent exacerbations, participants are re-randomised to the active treatment options.

# Model

The approach assumes a two-state transition model with participants moving between a healthy state and exacerbation state and where we can represent and capture the day each state begins.

The analysis adopts a discrete-time survival model the so-called *proportional continuation ratio model* which has the form of a logistic regression:

$$
\begin{aligned}
\log\left( \frac{\lambda(t|x)}{1-\lambda(t|x)} \right) &= \theta_{0t} + x' \beta
\end{aligned}
$$

where we take it that $x$ contains unit level covariates for baseline ppfev1, treatment assignment and subject level intercept to account for repeat measures, @Tutz2016.
The intercepts, $\theta_{0t}$, which vary over time, can be interpreted as the baseline hazard and the effect of the (time-constant) frailty is to provide an individual-level shift this baseline hazard.
The same model is adopted for both processes (each having their own set of parameters with the random effects being treated as independent) and are implemented and fit within a single stan model.

We assume that treatment effects enter into the linear predictor in both the exacerbation and healthy state; the treatment is randomised at the start of the exacerbation period and we retain this allocation to characterise carry over effects in the following healthy state.
At the next exacerbation, a new treatment is randomised.
Given that participants enrol into the study in a healthy state, no covariate is included in the model for this first period.

# Treatment Effects

After fitting the model to the available data, we use the posterior draws to repeatedly simulate the health state trajectory of individuals from trial cohorts over a `r l_spec_1$decision_fu` days under a given treatment policy.
For each participant, we calculate the number of days in the exacerbation state and then compute the average in the cohort.
The posterior view on the mean number of exacerbation days under each treatment policy is then contrasted to give the expected difference in mean exacerbation days.

The implication of the modelling approach is that an effect of treatment can arise from impacts to either state in the transition process.
That is, any of an increased recovery time during exacerbations or increased rate of relapse in healthy state, or both, will induce treatment effects.

# Decision Processes

At each analysis we use the difference in mean exacerbation days to compute the probability of non-inferiority.

We adopt a **non-inferiority** margin of `r l_spec_1$dec_eh_delta_ni` days and require a `r l_spec_1$dec_thresh_ni` probability that the treatments arms (def and dis) are lower than the margin to stop the trial for non-inferiority.

Additionally, if the the treatments arms (def and dis) are greater than the NI margin with probability `r l_spec_1$dec_thresh_fut` then we stop future enrolment into the arm for **futility**.

The soc remains the reference arm throughout the study.

# Scenarios

@tbl-scen-01 provides a summary of the key parameters used to configure the scenarios.
Each scenario was simulated `r l_spec_1$nsim` times with a cohort of `r sum(l_spec_1$N_pt)` participants with each trial having up to `r length(l_spec_1$N_pt)` analyses after `r paste0(cumsum(l_spec_1$N_pt), collapse = ", ")` are enrolled.

```{r, echo = F, eval = T}
#| label: tbl-scen-01
#| tbl-cap: 'Scenarios - key simulation parameters'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

i <- 1
d_scen <- data.table()
g_scen_names <- c()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  d_tmp <- data.table(
    desc = l_cfg$desc,
    b_trt_eh = sprintf("%.2f, %.2f", l_cfg$b_trt_eh[2], l_cfg$b_trt_eh[3]),
    b_trt_he = sprintf("%.2f, %.2f", l_cfg$b_trt_he[2], l_cfg$b_trt_he[3]),
    
    ni_margin = l_cfg$dec_eh_delta_ni,
    dec_thresh_ni = l_cfg$dec_thresh_ni,
    dec_thresh_fut = l_cfg$dec_thresh_fut,
    decision_fu = l_cfg$decision_fu
    )
  
  d_scen <- rbind(d_scen, d_tmp)
}



kableExtra::kbl(
  d_scen, 
  escape = TRUE, booktabs = TRUE, align = c("lcccccc"),
  linesep = "",
  col.names=c("Scenario", "Treatment effects (E)", "Treatment effects (H)", 
              "NI Margin", "NI Threshold", "Fut Threshold", "Dec FU Days"), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 9)  |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2:3, width = "2cm")|>
  kableExtra::column_spec(4:7, width = "1.5cm")
```


# Results

## Example Trial

```{r, echo = F}
l_cfg <- copy(l[[1]]$l_spec)
d_w <- copy(l[[1]]$d_w[sim == l_cfg$ex_trial_ix[1]])
```


The example is from the first scenario and takes the first simulation.
All the figures in this section relate to a single trial.

@fig-disc-eh-ex-1 shows the distribution of the number of exacerbations in the trial sample (which may be less than the maximum sample size due to early stopping and for this trial was `r sprintf("N = %.0f", d_w[, uniqueN(id)])`) over the `r l_spec_1$followup` day followup.

@fig-enrol-ex-1 shows the progression of participants entering the trial and @fig-enrol-ex-2 shows how each of the treatment arms grow as exacerbations occur.


```{r}
#| label: fig-disc-eh-ex-1
#| fig-cap: 'Distribution of exacerbations'
#| fig-height: 4
#| fig-width: 4
#| out-height: 4in
#| out-width: 4in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

# Distribution of exacerbations in this sample.


# number of unique elements in vec
d_fig <- d_w[, .(
  n_E = uniqueN(per_id[state == "E"])
  ), keyby = .(sim, id)
]

# d_fig[, .N, keyby = .(desc, sim, n_E)]

# https://github.com/tidyverse/ggplot2/issues/2051
ggplot(d_fig, aes(x = n_E)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_x_continuous("Exacerbations", breaks = 0:10) +
  scale_y_continuous("Proportion of Pts", 
                     breaks = seq(0, 1, by = 0.1))
```

```{r}
#| label: fig-enrol-ex-1
#| fig-cap: 'Participant entry into study'
#| fig-height: 4
#| fig-width: 4
#| out-height: 4in
#| out-width: 4in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)
d_w <- copy(l[[1]]$d_w[sim == l_cfg$ex_trial_ix[1]])


d_fig <- d_w[state == "H" & per_id == 1, .SD[1], keyby = id]
d_fig[, t0 := round(t0)]
d_fig <- d_fig[, .N, keyby = .(t0)]
d_fig[, cN := cumsum(N)]

ggplot(d_fig, aes(x = t0, y = cN)) +
  geom_step() +
  scale_x_continuous("Day", breaks = seq(0, max(d_fig$t0), by = 100)) +
  scale_y_continuous("N", breaks = seq(0, max(d_fig$cN), by = 100))
```


```{r}
#| label: fig-enrol-ex-2
#| fig-cap: 'Allocation to each treatment over enrolment period'
#| fig-height: 4
#| fig-width: 4
#| out-height: 4in
#| out-width: 4in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)
d_w <- copy(l[[1]]$d_w[sim == l_cfg$ex_trial_ix[1]])

d_enrol <- copy(l[[1]]$d_enrol[sim == l_cfg$ex_trial_ix[1]])

d_fig <- d_w[state == "E", .SD[1], keyby = .(id, per_id)]
# d_fig[, .N, by = trt]
d_fig[, ttrt := round(t0 + tstart)]
d_fig <- d_fig[, .N, keyby = .(ttrt, trt)]
d_fig[, cN := cumsum(N), keyby = trt]
d_fig <- base::merge(
  d_fig, CJ(ttrt = 0:max(d_fig$ttrt), trt = l_spec_1$trt_lab),
  by = c("ttrt", "trt"), all.y = T)
d_fig[, cN := nafill(cN, type = "locf"), keyby = trt]
d_fig[is.na(cN), cN := 0]
d_fig[, trt := factor(trt, levels = l_spec_1$trt_lab)]

# ix_id_smpl <- sample(unique(d_w$id), size = 10)
# ggplot(d_fig[id %in% ix_id_smpl], 
#        aes(x = ttrt, y = trt, group = id)) +
#   geom_line()

ggplot(d_fig, aes(x = ttrt, y = cN, group = trt, col = trt)) +
  geom_step() +
  scale_color_manual("", values = c(
    "#A85F18", "#CC8922", "#CC6B22", "#228DCC"
  )) +
  scale_x_continuous("Day", breaks = seq(0, max(d_fig$ttrt), by = 200)) +
  scale_y_continuous("N", breaks = seq(0, max(d_fig$cN), by = 100))
```

@fig-post-par-ex-1 shows the prior and posterior distribution on the parameters from the model at the time of the last analysis.
These include the intercept terms for the piecewise baseline hazard, the covariate effects for baseline ppfev1, treatment effects and variance components used to account for repeat measures.

@fig-post-soj-ex-1 shows the prior and posterior on the sojourn and difference in sojourn times in each state (**H**ealthy and **E**xacerbation).
As noted earlier, to derive these, we use the model joint posterior to simulate the trajectory for a set of participants. 
Key points in this process are:

1. We simulate a trial population using the same number as presently observed in the analysis for each draw from the model parameters estimates
2. All participants are assumed to start in an exacerbation state and their state is updated by day
3. The decision horizon is set to `r l_spec_1$decision_fu` days
4. We assume a fixed treatment policy over the decision horizon and the baseline covariates are resampled from the actual trial population
5. For each participant in the population, we compute the total days in an exacerbation state
6. For each simulated population we compute the mean days in exacerbation state

Point (4) makes the strong assumption that while carry over effects of a policy may exist, their is no cumulative impact beyond that.
Point (6) gives us a distribution of mean days in exacerbation state with the uncertainty propagated from the posterior.

The steps are repeated for each treatment arm and the results to create treatment comparisons.

The equivalent process can be run to obtain a view on the prior for these parameters.

@tbl-soj-ex-01 summarises the distribution of mean days in exacerbation and health state under each treatment arm for this specific example.
@tbl-diff-ex-01 summarises the difference between each treatment arm and soc.

```{r}
#| label: fig-post-par-ex-1
#| fig-cap: 'Prior and posterior for model parameters'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)


d_pri_par <- copy(l[[1]]$d_pri_par[sim == l_cfg$ex_trial_ix[1]])
d_post_par <- copy(l[[1]]$d_post_par[sim == l_cfg$ex_trial_ix[1]])

# only get the last one:

d_pri_par <- d_pri_par[ic == max(ic)]
d_post_par <- d_post_par[ic == max(ic)]

d_pri_par[, `:=`(ic = NULL)]
d_post_par[, `:=`(ic = NULL)]

d_pri_par[, desc := "prior"]
d_post_par[, desc := "posterior"]

d_par <- rbind(d_pri_par, d_post_par)

d_fig <- melt(d_par, id.vars = c("desc", "sim"))
d_fig[, desc := factor(desc, levels = c("posterior", "prior"))]

p_1 <- ggplot(d_fig[variable %like% "a_"], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_x_continuous("") +
  scale_linetype_discrete("") +
  ggh4x::facet_wrap2(~variable, scales = "free", ncol = 3) +
  ggh4x::facetted_pos_scales(
    x = list(
      variable %like% "a_he" ~ scale_x_continuous(limits = c(-8, 0)),
      variable %like% "a_eh" ~ scale_x_continuous(limits = c(-5, 1))
    )
  )

p_2 <- ggplot(d_fig[!(variable %like% "a_")], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 4)

p_1 / p_2 +
  plot_layout(heights = c(1, 1), guides = "collect")

```


```{r}
#| label: fig-post-soj-ex-1
#| fig-cap: 'Prior and posterior for sojourn times'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

l_cfg <- copy(l[[1]]$l_spec)

d_pri_res <- copy(l[[1]]$d_pri_res[sim == l_cfg$ex_trial_ix[1]])
d_post_res <- copy(l[[1]]$d_post_res[sim == l_cfg$ex_trial_ix[1]])

# only get the last one:

d_pri_res <- d_pri_res[ic == max(ic)]
d_post_res <- d_post_res[ic == max(ic)]

d_pri_res[, `:=`(ic = NULL)]
d_post_res[, `:=`(ic = NULL)]

d_pri_res[, desc := "prior"]
d_post_res[, desc := "posterior"]

d_par <- rbind(d_pri_res, d_post_res)

d_fig <- melt(d_par, id.vars = c("desc", "sim"))
d_fig[, desc := factor(desc, levels = c("posterior", "prior"))]

p_1 <- ggplot(d_fig[variable %like% "mu_"], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  ggh4x::facetted_pos_scales(
    x = list(
      variable %in% c("mu_soc_H", "mu_def_h", "mu_dis_H") ~ scale_x_continuous(limits = c(50, 100)),
      variable %like% "_E" ~ scale_x_continuous(limits = c(0, 25))
    )
  )

p_2 <- ggplot(d_fig[!(variable %like% "mu_")], 
              aes(x = value, group = desc, colour = desc, lty = desc)) +
  geom_density() +
  scale_colour_discrete("") +
  scale_linetype_discrete("") +
  facet_wrap(~variable, scales = "free", ncol = 2) +
  ggh4x::facetted_pos_scales(
    x = list(
      variable %in% c("del") ~ scale_x_continuous(limits = c(-10, 10))
    )
  )

p_1 / p_2 +
  plot_layout(heights = c(1, 1), guides = "collect")
```


```{r, echo = F}

l_cfg <- copy(l[[1]]$l_spec)
d_trt <- copy(l[[1]]$d_trt_effects[sim == l_cfg$ex_trial_ix[1]])
d_pr_dec <- copy(l[[1]]$d_pr_dec[sim == l_cfg$ex_trial_ix[1]])

d_pr_dec <- dcast(d_pr_dec, sim + ic + trt ~ rule, value.var = c("p"))
setnames(d_pr_dec, "fut", "pr_fut")
setnames(d_pr_dec, "ni", "pr_ni")

d_trt[, N := cumsum(l_cfg$N_pt)[ic]]
d_trt[, desc := l_cfg$desc]

d_trt[par %like% "_E", state := "Exacerbation"]
d_trt[par %like% "_H", state := "Healthy"]
d_trt[par %like% "_def_", trt := "def"]
d_trt[par %like% "_dis_", trt := "dis"]
d_trt[par %like% "_soc_", trt := "soc"]

d_trt[par %like% "del_",  stat := "diff"]
d_trt[par %like% "mu_",  stat := "mean"]

d_trt[state == "Exacerbation", out := sprintf("%.1f (%.1f, %.1f)", mu, lwr, upr)]
d_trt[state == "Healthy", out := sprintf("%.1f (%.1f, %.1f)", mu, lwr, upr)]


d_trt <- base::merge(d_trt, d_pr_dec, by = c("sim", "ic", "trt"), all.x = T)

```

```{r, echo = F, eval = T}
#| label: tbl-soj-ex-01
#| tbl-cap: 'Posterior mean and 95% CrI sojourn times'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H


d_trt[stat == "mean" & out == "NA (NA, NA)", out := NA]

d_tbl <- dcast(d_trt[stat == "mean"], desc + state + trt ~ N, value.var = "out")


options(knitr.kable.NA = "-")

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 9) 
```


```{r, echo = F, eval = T}
#| label: tbl-diff-ex-01
#| tbl-cap: 'Posterior difference in sojourn times'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

d_tbl <- d_trt[stat == "diff"]
d_tbl <- d_tbl[order(trt, N)]
d_tbl[out == "NA (NA, NA)", out := NA]

options(knitr.kable.NA = "-")

kableExtra::kbl(
  d_tbl[state == "Exacerbation", .(desc, N, trt, out, pr_fut, pr_ni)], 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "\\addlinespace"),
  col.names=c("Scenario", "N", "Treatment", "Differnce", "Pr(fut)", "Pr(NI)"), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 9) 
```


## Trial Population

```{r, eval = T}
#| label: exac-comp
#| code-summary: Frequency of exacerbation
#| echo: FALSE

i <- 1
d_w <- data.table()

g_scen_names <- c()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # add desc...
  d_tmp <- copy(l[[i]]$d_w)
  d_tmp[, desc := l_cfg$desc]
  
  g_scen_names <- c(g_scen_names, l_cfg$desc)
  
  d_tmp[, ix_desc := i]
  d_w <- rbind(d_w, d_tmp)
}

d_w[, desc := factor(desc, levels = g_scen_names)]
d_w[, len_seg := tstop - tstart]


# Pick out trial data from one simulation for each scenario
ix_smpl <- 1:5 # sort(sample(1:max(d_w$sim), 5))
d_w_smpl <- d_w[sim %in% ix_smpl, ]
```

@fig-dist-n-eh shows the distribution of number of exacerbations for a random selection of simulations and @fig-dist-dur-eh shows the distribution of exacerbation days over all exacerbations that occurred.

@fig-dist-dur_he shows the distribution of health days for the first healthy period (all participants are assumed to enter in a healthy state and may progress to one or more exacerbations).
@fig-dist-n-trt shows the distribution of treatment allocations over the entire follow up period.
@fig-dist-n-enrol shows the distribution of enrolment over the all simulations for each scenario.

```{r}
#| label: fig-dist-n-eh
#| fig-cap: 'Number of exacerbations in trial population over 365 days followup for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[, .(
  n_E = uniqueN(per_id[state == "E"])
  ), keyby = .(desc, sim, id)
]

# d_fig[, .N, keyby = .(desc, sim, n_E)]

# https://github.com/tidyverse/ggplot2/issues/2051
ggplot(d_fig, aes(x = n_E)) +
  geom_bar(aes(y = after_stat(prop), group = 1)) +
  scale_x_continuous("Exacerbations", breaks = 0:10) +
  scale_y_continuous("Proportion of Pts", 
                     breaks = seq(0, 1, by = 0.1)) +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl),
             labeller = label_wrap_gen(width = 20))

```


```{r}
#| label: fig-dist-dur-eh
#| fig-cap: 'Duration of exacerbations in trial population over 365 days followup for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "E", .(dur_eh = sum(len_seg)), keyby = .(desc, sim, id, per_id)]

ggplot(d_fig, aes(x = dur_eh)) +
  geom_histogram(bins = 20) +
  scale_x_continuous("Duration (days)") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl),
             labeller = label_wrap_gen(width = 20))

```


```{r}
#| label: fig-dist-dur_he
#| fig-cap: 'Duration of first healthy period in trial population for random subset of simulations'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "H" & per_id == 1, .(dur_he_1 = sum(len_seg)), keyby = .(desc, sim, id)]

ggplot(d_fig, aes(x = dur_he_1)) +
  geom_histogram(bins = 20) +
  scale_x_continuous("Duration (days)") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl),
             labeller = label_wrap_gen(width = 20))

```


```{r}
#| label: fig-dist-n-trt
#| fig-cap: 'Distribution of treatment allocation (randomisation of each treatment arm over trial population)'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

d_fig <- d_w_smpl[
  state == "E", .N, keyby = .(desc, sim, id, per_id, trt)]
d_fig[, `:=`(N = NULL)]
d_fig <- d_fig[, .N, keyby = .(desc, sim, trt)]
d_fig[, trt := factor(trt, levels = l_spec_1$trt_lab)]
# d_fig[, sum(N), by = sim]

ggplot(d_fig, aes(x = trt, y = N)) +
  geom_bar(stat = "identity") +
  scale_x_discrete("") +
  scale_y_continuous("N") +
  facet_wrap(desc ~ paste0("sim: ", sim), ncol = length(ix_smpl),
             labeller = label_wrap_gen(width = 20))

```



```{r}
#| label: fig-dist-n-enrol
#| fig-cap: 'Distribution of enrolment size over all simulations'
#| fig-height: 3
#| fig-width: 6
#| out-height: 3in
#| out-width: 6in
#| fig-dpi: 300
#| fig-align: "center"
#| fig-pos: H

# unique pt enrolled by sim
d_fig <- d_w[, .(N = length(unique(id))), keyby = .(desc, sim)]
d_fig <- d_fig[, .(freq = .N), keyby = .(desc, N)]

# d_fig[, sum(N), by = sim]

ggplot(d_fig, aes(x = N, y = freq)) +
  geom_bar(stat = "identity", width = 10) +
  scale_x_continuous("Enrolment", breaks = cumsum(l_spec_1$N_pt)) +
  scale_y_continuous("N") +
  facet_wrap(desc ~ ., nrow = 1)
```

## Parameters 

@tbl-par-comp summarises the model parameter estimates, by scenario and analysis, averged over the simulations.

```{r, eval = T}
#| label: tbl-par-comp
#| tbl-cap: 'Posterior parameter summary'

i <- 1
d_par <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_tmp <- copy(l[[i]]$d_post_smry_1[
    par %in% l_spec_1$par_names])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_tmp[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  d_smry <- d_tmp[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_par <- rbind(
    d_par,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_par[, stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]

d_tbl <- dcast(d_par, desc + par ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl <- d_tbl[order(desc, par)]

kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llccc"), longtable = TRUE,
  col.names=c("Scenario", "Parameter", cumsum(l_cfg$N_pt)), digits = 1)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3:5, width = "2.5cm") 

```


{{< pagebreak >}}

## Sojourn Time

@tbl-soj-01 shows the expectations and 95% uncertainty interval for the posterior means of the sojourn times (days) in each health state by treatment arm over a simulated `r l_spec_1$decision_fu` day decision horizon.

@tbl-soj-02 provides the difference in sojourn times by health state in each treatment arm relative to the control arm.

```{r, echo = F, eval = T}
i <- 1
d_soj <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_par <- copy(
    l[[i]]$d_trt_effects[par %in% c(
      "mu_soc_H", "mu_def_H", "mu_dis_H",
      "mu_soc_E", "mu_def_E", "mu_dis_E"
      )])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_par[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  # only summarising the means
  d_smry <- d_par[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_soj <- rbind(
    d_soj,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_soj[par %like% "_E", state := "Exacerbation"]
d_soj[par %like% "_H", state := "Healthy"]
d_soj[state == "Exacerbation", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]
d_soj[state == "Healthy", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]

d_soj[par %like% "_def_", trt := "def"]
d_soj[par %like% "_dis_", trt := "dis"]
d_soj[par %like% "_soc_", trt := "soc"]

d_tbl <- dcast(d_soj, desc + state + trt ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, trt := factor(trt, levels = l_spec_1$trt_lab)]
d_tbl[, state := factor(state, levels = c("Healthy", "Exacerbation"))]
d_tbl <- d_tbl[order(desc, trt, state)]
```


```{r, echo = F, eval = T}
#| label: tbl-soj-01
#| tbl-cap: 'Expectation of posterior mean sojourn times by scenario, health state and treatment arm'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H


options(knitr.kable.NA = "-")

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "",  "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 1)  |>
  kableExtra::kable_styling(font_size = 9)  |>
  kableExtra::column_spec(1, width = "3.5cm")
```

```{r, echo = F, eval = T}
i <- 1
d_soj <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_par <- copy(
    l[[i]]$d_trt_effects[par %in% c(
      "del_def_H", "del_dis_H", 
      "del_def_E", "del_dis_E"
      )])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate the rmst forward to impute the effects in those sims that had
  # early stopping
  d_par[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  # only summarising the means
  d_smry <- d_par[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_soj <- rbind(
    d_soj,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
}

d_soj[par %like% "_E", state := "Exacerbation"]
d_soj[par %like% "_H", state := "Healthy"]
d_soj[state == "Exacerbation", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]
d_soj[state == "Healthy", stat := sprintf("%.1f (%.1f, %.1f)", mu, lo, hi)]

d_soj[par %like% "_def_", trt := "def"]
d_soj[par %like% "_dis_", trt := "dis"]
d_soj[par %like% "_soc_", trt := "soc"]

d_tbl <- dcast(d_soj, desc + state + trt ~ N, value.var = "stat")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, trt := factor(trt, levels = l_spec_1$trt_lab)]
d_tbl[, state := factor(state, levels = c("Healthy", "Exacerbation"))]
d_tbl <- d_tbl[order(desc, trt, state)]
```


```{r, echo = F, eval = T}
#| label: tbl-soj-02
#| tbl-cap: 'Expectation of difference in mean sojourn times by scenario, health state and treatment arm relative to the control arm'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "\\addlinespace"),
  col.names=c("Scenario", "State", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 1)  |>
  kableExtra::kable_styling(font_size = 9)   |>
  kableExtra::column_spec(1, width = "3.5cm")
```



{{< pagebreak >}}


## Decision Probability

@tbl-dec-pr-1 shows the cumulative probability of the non-inferiority and futility decisions based on the difference in the exacerbation sojourn times.

```{r, echo = F, eval = T}
#| label: tbl-dec-pr-1
#| tbl-cap: 'Cumulative probability of each decision type'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H


i <- 1
d_dec <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_pr_dec[trt %in% c("def", "dis")])
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  
  # number of enrolments at each interim (interim sample size sequence)
  d_N <- data.table(ic = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  d_dec_1[, cdec := as.integer(cumsum(dec)>0), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := nafill(cdec, type = "locf"), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := as.logical(cdec)]
  
  d_dec_1 <- base::merge(d_dec_1, d_N, by = "ic")
  
  # cumulative proportion for which each decision quantity has been met by 
  # analysis and domain
  d_dec_1 <- d_dec_1[, .(pr_val = mean(cdec)), keyby = .(ic, N, rule, trt)]
  
  d_dec <- rbind(
    d_dec,
    cbind(scenario = i, desc = l_cfg$desc, d_dec_1)
  )

}

d_tbl <- dcast(
  d_dec, desc + rule  + trt ~ N, value.var = "pr_val")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, rule := factor(rule, levels = c("ni", "fut"), labels = c("NI", "Fut"))]
d_tbl <- d_tbl[order(desc, rule)]


kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "\\addlinespace"),
  col.names=c("Scenario", "Rule", "Treatment", cumsum(l_cfg$N_pt)), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 8) |>
  kableExtra::column_spec(1, width = "5cm")|>
  kableExtra::column_spec(2:6, width = "1.5cm")



```





{{< pagebreak >}}

## Sample Size

Expected number of participant randomisations on each treatment arm under each scenario.

```{r, echo = F, eval = F}
#| label: sample-size-comp
#| code-summary: Compute expected sample size


i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  # l_cfg$desc
  
  N_analysis <- length(l_cfg$N_pt)
  N_trt <- 3
  # 
  d_grid <- CJ(
    sim = 1:l_cfg$nsim,
    ic = 1:N_analysis,
    trt = 1:N_trt
  )
  d_grid$trt <- l[[i]]$l_spec$trt_lab[d_grid$trt]
    
  d_all <- copy(l[[i]]$d_all[state == "E"])
  
  # time of last enrolment entering into interim analysis by sim
  d_analy_set <- copy(l[[i]]$d_analys_sets)
  d_analy_set <- d_analy_set[!is.na(t0_last), ]
  d_analy_set[, ic := 1]
  d_analy_set[, ic := cumsum(ic), by = sim]
  
  setkey(d_all, sim, t0)
  setkey(d_analy_set,  sim, t0_last)

  
  # identify the analysis that each pt first enters
  # Rolling join: find first cutoff_date >= enrol_date
  d_all[
    d_analy_set,
    on = .(sim, t0 = t0_last),
    roll = -Inf,
    ic := i.ic
  ]
  # last one carried back
  d_all[, ic := nafill(ic, type = "nocb"), by = sim]
  
  # how many randomisations per interim and trt arm (pt may have more than 
  # one randomisation over their followup)
  d_smry <- base::merge(
    d_grid,
    d_all[, .(.N), keyby = .(sim, ic, trt)],
    by = c("sim", "ic", "trt"),
    all.x = T
  )
  
  # some arms may stop after interim 1 and so there would be no count for 
  # subsequent interims
  d_smry[is.na(N), N := 0]
  d_smry[, N_cmtv := cumsum(N), keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_smry
    )
  )
  
}

d_N_trt <- d_N_by_rand[, .SD[.N], keyby = .(scenario, desc, sim, trt)]
d_N_tot <- d_N_trt[, .(N_cmtv = sum(N_cmtv)), keyby = .(scenario, desc, sim) ]

# add in total sample size
d_tbl <- rbind(
  d_N_trt[, .(N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc, trt)],
  d_N_tot[, .(trt = "all", N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc)]
)
setkey(d_tbl, scenario, desc)


kableExtra::kbl(
  d_tbl[, .SD, .SDcols = -c("scenario")], 
  col.names = c("Desc", "TRT", "Mean", "SD"),
  format = "simple", digits = 1)

```



{{< pagebreak >}}
# References {.unnumbered}

<!-- Needs to have a citation for this to work otherwise you will get the old \end{CSLReferences} error  -->

::: {#refs}
:::

{{< pagebreak >}}




