---
title: BEAT CF MARS
subtitle: |
    Simulation Report
date: last-modified
date-format: "D MMMM YYYY"
author: Mark Jones
editor: source
toc: true
toc-depth: 3
bibliography: ../etc/refs.bib
csl: ../etc/elsevier-harvard.csl
include-in-header:
  text: |
    \usepackage{makecell}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r}
#| label: load-data
#| code-summary: Data load
#| echo: false

# uml digs
suppressPackageStartupMessages(library(nomnoml))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(qs))
suppressPackageStartupMessages(library(git2r))
suppressPackageStartupMessages(suppressWarnings(library(gt)))
suppressPackageStartupMessages(library(ggh4x))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(pbapply))


toks <- unlist(tstrsplit(getwd(), "/")) 
if(toks[length(toks)] == "beatcf-mars-sim"){
  prefix_cfg <- "./etc/sim18/"
  prefix_stan <- "./stan"
  prefix_fig <- "./fig"
  prefix_data <- "./data"
  prefix_r <- "./R"
} else {
  prefix_cfg <- "../etc/sim18/"
  prefix_stan <- "../stan"
  prefix_fig <- "../fig"
  prefix_data <- "../data"
  prefix_r <- "../R"
}



source(paste0(prefix_r, "/libs.R"))
source(paste0(prefix_r, "/init.R"))
source(paste0(prefix_r, "/util.R"))
source(paste0(prefix_r, "/data-sim18.R"))


# Each input file corresponds to the results from a single simulation
# scenario/configuration.
# Load all the files into a single list.

# files of interest
sim_lab <- "sim18-04"

flist <- list.files(paste0(prefix_data, "/", sim_lab), pattern = "sim18")
toks <- list()
l <- list()
i <- 1

for(i in 1:length(flist)){

  l[[i]] <- qs::qread(file.path(paste0(prefix_data, "/", sim_lab), flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

N_sims <- l[[1]]$cfg$nsim



# temp to replace
# l_spec <- get_prototype_cfg()
l_spec_1 <- l[[1]]$l_spec



desc_labs <- character()

for(i in 1:length(l)){
  
  desc_labs <- c(desc_labs, l[[i]]$l_spec$desc)
  
}

```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  BEAT-CF MARS |
|Reference intervention: |  Standard airway clearance therapy at least twice per day for 14 days, plus 14 days of oral antibiotics commencing as soon as practicable (and ≥ 72 hours and < 7 days) after onset of new or acute worsening of cough. |   
|Intervention: | (1) Deferred antibiotic strategy (2) Early discontinuation antibiotic strategy. |   
|Outcome: | Time to recovery from onset of exacerbation |   
|Study design:  |   Bayesian adaptive trial with early stopping rules | 
|Sponsor:  |    University of Sydney, NSW, Australia | 
Protocol: |  todo |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Tom Snelling | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}


\tiny
| Version    |   Date     | Change    |   Reason     |
|:----|:----|:--------|:--------|
| 0.1 | 2026-03-04 | First version revised design | - |
| 0.2 | 2026-03-04 | Incorporate carry-over effects into the healthy state | Feedback from CI |
| 0.3 | 2026-05-18 | Discrete time process | Better representation of state |
| 0.4 | 2026-06-03 | Revise follow up and use direct parameters for treatment effects | Simplify design |
| 0.5 | 2026-06-23 | Same idea as before but align the model with longitundinal ordinal | More realistic model |
\normalsize


{{< pagebreak >}}

# Design

Beat CF Mars is a three arm, sequential trial with adaptive stopping rules based on non-inferiority and futility, see @Berry2011.
The primary endpoint is the symptom free days to `r l_spec_1$followup_dec` days.

Participants are enrolled in a well state and monitored over `r l_spec_1$followup` days.
On exerperiencing an exacerbation in the monitoring period, participants are randomised to:

+ standard of care antibiotics (soc)
+ deferred antibiotics (def)
+ early discontinue antibiotics (dis)

and tracked over `r l_spec_1$followup_dec` days for resolution of symptoms and/or the occurrence of subsequent exacerbations, the goal being to observe the number of symptom free days over the `r l_spec_1$followup_dec` window.

Once an exacerbation occurs, we follow participants over a full `r l_spec_1$followup_dec` day window irrespective of whether this overflows the `r l_spec_1$followup` day monitoring period.

If a participant doesn't have an exacerbation, then they would not receive any randomised treatment and so they would not contribute to the analyses.
Therefore, what we would be characterising amounts to *"Among CF patients who experience an exacerbation that requires treatment, what is the difference in symptom free days in alternative antibiotic strategies over a pre-specified follow up period?"*

If a participant has a subsequent exacerbation within the `r l_spec_1$followup` monitoring window, they are re-randomised and the process repeats.

The simulation label for this report is `r sim_lab` and `r l_spec_1$nsim` trials were simulated for each scenario.

# Model

A longitudinal markov ordinal model is used to fit individual level data over collected by surveys at days: `r paste0(l_spec_1$visit_days, collapse = ", ")`.

The model incorporates intercepts that reflect the cutpoints along with parameters to account for:

+ treatment assignment
+ symptoms at previous observation
+ time (quadratic)
+ treatment by time
+ previous state by time

A markov assumption accounts for the repeat measure over time.

Variation induced via a previous state by treatment could be considered for the model.
The latter modifies tretament efficacy dependent on the patient's disease state.

A participant level random effect would need to be included to account for re-randomisation.

# Treatment Effects

Treatment effects represent the difference in expected symptom free days to `r l_spec_1$followup_dec` across groups.

# Decision Processes

At each analysis we use the difference in mean exacerbation days to compute the probability of non-inferiority.

**Non-inferiority:**

```{r, echo = F}
ni_delta <- l_spec_1$dec_delta_ni
threshold_ni <- l_spec_1$dec_thresh_ni
threshold_fut <- l_spec_1$dec_thresh_fut
```


It is reasonable to suppose symptom free days could be lower in the treatment groups than under standard of care, that is the number of symptom free days is reduced by virtue of a deferred or early discontinuation strategy relative to soc. 
We define a non-inferiority margin to accept up to `r l_spec_1$dec_delta_ni` day reduction in symptom free days. 
We defined non-inferiority if:

$$
\begin{aligned}
Pr(\text{days}_{trt} - \text{days}_{soc} > -`r ni_delta`) > `r threshold_ni`
\end{aligned}
$$

such that the treatment would be interpreted to be non-inferior to standard of care.

**Futility:**

An analogous definition is assumed for futility:

$$
\begin{aligned}
Pr(\text{days}_{trt} - \text{days}_{soc} < -`r ni_delta`) > `r threshold_fut`
\end{aligned}
$$

the treatment is considered non-inferior to standard of care.

Treatment arms may be stopped for NI or futility but the soc remains the reference arm throughout the study.

# Scenarios

@tbl-scen-01 provides a summary of the key parameters used to configure the scenarios.
Given the model includes interaction terms, the treatment effects represent the true difference in sojourn times with no symptoms that are induced by the combination of all the parameters within the linear predictor.

Each scenario was simulated `r l_spec_1$nsim` times with a cohort of `r sum(l_spec_1$N_pt)` participants with each trial having up to `r length(l_spec_1$N_pt)` analyses after `r paste0(cumsum(l_spec_1$N_pt), collapse = ", ")` are enrolled.
The final analysis is run on all participants, i.e. we assume that followup was completed.

The data generating process assumes a Markov progression of state.
We assume that data is collected over a subset of the entire followup.
Importantly, for the days with gaps above one day (e.g. day 21) we assume that both the day 20 and day 21 data on state are collected. 

```{r, echo = F, eval = T}
#| label: tbl-scen-01
#| tbl-cap: 'Scenarios - key simulation parameters'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

i <- 1
d_scen <- data.table()
g_scen_names <- c()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  d_tmp <- data.table(
    desc = l_cfg$desc,
    b_trt = sprintf("%.2f, %.2f", l_cfg$delta_dur_tru[1], l_cfg$delta_dur_tru[2]),
    
    ni_margin = l_cfg$dec_delta_ni,
    dec_thresh_ni = l_cfg$dec_thresh_ni,
    dec_thresh_fut = l_cfg$dec_thresh_fut,
    followup_dec = l_cfg$followup_dec
    )
  
  d_scen <- rbind(d_scen, d_tmp)
}



kableExtra::kbl(
  d_scen, 
  escape = TRUE, booktabs = TRUE, align = c("lccccc"),
  linesep = "",
  col.names=c("Scenario", "Treatment effects (diff in expected time with no symptom)", 
              "NI Margin", "NI Threshold", "Fut Threshold", "Dec FU Days"), 
  digits = 3)  |>
  kableExtra::kable_styling(font_size = 9)  |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "3cm")|>
  kableExtra::column_spec(3, width = "2cm")|>
  kableExtra::column_spec(4:6, width = "1.5cm")
```

Files used:

```{r}
for(i in 1:length(flist)){
  cat("File: ", flist[i], "\n")
}
```


Further configuration from first simulation scenario.

```{r, echo = F, eval = T}
#| label: tbl-cfg-1
#| tbl-cap: 'Scenarios 1 - selected config.'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

d_l_spec <- sim18_cfg_to_dtb(l_spec_1)

kableExtra::kbl(
  d_l_spec[], 
  escape = TRUE, booktabs = TRUE, align = c("lr"),
  linesep = "")  |>
  kableExtra::kable_styling(font_size = 7)  
```



# Results


```{r, echo = F}
ex_trial_plots <- function(id_scen = 1, id_ex = 1, include_tran_m = F){
  
  l_cfg <- copy(l[[id_scen]]$l_spec)
  stopifnot(id_ex <= l_cfg$nex)
  
  d_all <- copy(l[[id_scen]]$d_all[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_smry_1 <- copy(l[[id_scen]]$d_post_smry_1[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_smry_2 <- copy(l[[id_scen]]$d_post_smry_2[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_smry_3 <- copy(l[[id_scen]]$d_post_smry_3[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_smry_4 <- copy(l[[id_scen]]$d_post_smry_4[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_par <- copy(l[[id_scen]]$d_post_par[sim == l_cfg$ex_trial_ix[id_ex]])
  d_post_sop <- copy(l[[id_scen]]$d_post_sop[sim == l_cfg$ex_trial_ix[id_ex]])
  
  
  # observed data - proportion by day and state ------------
  d_fig_1 <- copy(d_all)
  d_fig_1[, state := factor(
    state, levels = l_cfg$state_opts, labels = names(l_cfg$state_opts))]
  d_fig_1[, trt := factor(trt, levels = l_cfg$trt_lab)]
  
  p_1 <- ggplot(d_fig_1, aes(fill = state, x = day)) +
    geom_bar(position = "fill") +
    scale_fill_discrete("") +
    scale_x_continuous("", breaks = 1:max(d_fig_1$day)) +
    facet_wrap(~trt, ncol = 1) +
    theme(
      legend.position = "bottom"
    )
 
  
  # Observed vs model sop -------------------------------
  # Observed data, filling in missing parts.
  d_fig_1 <- copy(d_all)
  d_fig_1[, state := factor(
    state, levels = l_cfg$state_opts, labels = names(l_cfg$state_opts))]
  d_fig_1[, trt := factor(trt, levels = l_cfg$trt_lab)]
  
  d_fig_2 <- d_fig_1[, .(.N), keyby = .(state, trt, day)]
  d_fig_2 <- base::merge(
    d_fig_2, 
    d_fig_1[, .(N_unit = length(unique(id))), keyby = .(trt)],
    by = "trt", all.x = T
  )
  d_fig_2[, prop := N/N_unit]
  d_fig_2 <- base::merge(
    CJ(trt = l_spec_1$trt_lab, 
       state = names(l_spec_1$state_opts), 
       day = l_spec_1$visit_days[-1]), 
    d_fig_2, 
    by = c("trt", "state", "day"),
    all.x = T
  )
  d_fig_2[is.na(prop), prop := 0]
  d_fig_2[, source := "Observed"]
  d_fig_2[, `:=`(N = NULL, N_unit = NULL)]
  
  stopifnot(
    nrow(d_fig_2) == 
      length(l_cfg$state_lab) * 
      length(l_cfg$trt_lab) * 
      length(l_cfg$visit_days[-1]))
    
  # Summary from model parameters
  d_fig_3 <- d_post_smry_2[ic == d_post_smry_2[!is.na(mu), max(ic)]]
  d_fig_3 <- d_fig_3[, .(trt, state, day, prop = mu, lo, hi, source = "Model")]
  
  
  d_fig <- rbind(d_fig_2, d_fig_3, fill = T)
  d_fig[, state := factor(state, levels = l_cfg$state_lab)]
  d_fig[, trt := factor(trt, levels = l_cfg$trt_lab)]
    
  p_2 <- ggplot(
    d_fig[source == "Model"], aes(x = day, y = prop, group = trt)) +
    # geom_linerange(aes(ymin = lo, ymax = hi), lwd = 0.3) +
    geom_line(aes(lty = trt), lwd = 0.3) +
    geom_point(
      data =  d_fig[source == "Observed"], 
      aes(x = day, y = prop, col = trt),
      inherit.aes = F, shape = 1, size = 0.89
    ) +
    scale_x_continuous("", breaks = seq(1, max(d_fig_1$day), by = 2)) +
    scale_y_continuous("SOP", breaks = seq(0, 1, by = 0.1)) +
    scale_color_discrete("") +
    scale_linetype_discrete("") +
    facet_wrap(. ~ state, ncol = 3) +
    theme(
      legend.position = "bottom"
    )
  
  
  p_3 <- ggplot(
    d_fig[source == "Model"], aes(
      x = day, y = prop, group = trt, col = trt, lty = trt)) +
    geom_line(lwd = 0.3) +
    scale_x_continuous("", breaks = seq(1, max(d_fig_1$day), by = 2)) +
    scale_y_continuous("SOP", breaks = seq(0, 1, by = 0.1)) +
    scale_color_discrete("") +
    scale_linetype_discrete("") +
    facet_wrap(. ~ state, ncol = 3) +
    theme(
      legend.position = "bottom"
    )
  
  # transition matrix ------------------
  d_tran <- data.table()
  if(include_tran_m){
    m_post <- as.matrix(d_post_par)
  l_spec_mod <- copy(l_cfg)
  t_days <- 1:max(l_cfg$visit_days)
  t_gaps <- rep(1, length = max(l_cfg$visit_days))
  # gaps are indexing
  t_gaps[t_gaps == 7] <- 2
  ii <- jj <- kk <- 1
  
  d_tran <- rbindlist(lapply(1:nrow(m_post), function(ii){
    
    # calculate sop based on parameter estimates at their post mean
    l_spec_mod$alpha <- as.numeric(m_post[ii , c("a[1]", "a[2]")])
    l_spec_mod$b_trt <- as.numeric(
      m_post[ii , c("b_trt[1]", "b_trt[2]", "b_trt[3]")])  
    l_spec_mod$b_prev <- as.numeric(
      m_post[ii , c("b_prev[1]", "b_prev[2]", "b_prev[3]")])  
    l_spec_mod$b_time_1 <- as.numeric(m_post[ii , c("b_time_1")])  
    l_spec_mod$b_time_2 <- as.numeric(m_post[ii , c("b_time_2")]) 
    l_spec_mod$b_trt_time <- as.numeric(
      m_post[ii , c("b_trt_time[1]", "b_trt_time[2]", "b_trt_time[3]")])  

    names(l_spec_mod$b_trt) <- l_spec_mod$trt_lab
    names(l_spec_mod$b_trt_time) <- l_spec_mod$trt_lab
      
    d_tran_trt <- data.table()
    for(jj in seq_along(l_cfg$trt_lab)){
      for(kk in seq_along(t_days)){
      
        
        d_pr <- data.table(
          sim18_transition_matrix(
            t_days[kk], l_cfg$trt_lab[jj], l_spec_mod)
          )
        d_pr[, `:=`(
          from = l_spec_mod$state_lab, 
          trt = l_cfg$trt_lab[jj],
          day = t_days[kk])]
        d_pr <- melt(
          d_pr, id.vars = c("trt", "from", "day"), 
          variable.name = "to", value.name = "p_tran"
        )
        d_tran_trt <- rbind(
          d_tran_trt, d_pr
        )
      }
    }
    d_tran_trt
    
    }), idcol = "id_draw" )
  
    d_tran <- dcast(d_tran[, .(
      mu = mean(p_tran)
    ), keyby = .(day, trt, from, to)], day + trt + from ~ to, value.var = "mu")
    d_tran[, from := factor(from, levels = l_cfg$state_lab)]
    setkey(d_tran, day, trt, from)
  }
  
  
  # durations and trt effects -------------
  d_post_sop[, gap := day - data.table::shift(day, type = "lag"), 
             keyby = .(id_draw, trt, ic)]

  # posterior days in each state
  d_dur <- melt(
    d_post_sop[day > 0 & ic == max(d_post_sop$ic), .(
      none = sum(none),
      mild = sum(mild),
      severe = sum(severe)
      ), keyby = .(trt, id_draw)], id.vars = c("trt", "id_draw"), 
    variable.name = "state", value.name = "days")
  
  d_fig_1 <- dcast(d_dur, state + id_draw ~ trt, value.var = "days")
  d_fig_1[, delta_def := def - soc]
  d_fig_1[, delta_dis := dis - soc]
    
  d_fig_2 <- melt(d_fig_1, id.vars = c("state", "id_draw"), 
                  variable.name = "par")
    
  p_4 <- ggplot(
    d_fig_2[!(par %like% "delta")], aes(x = value, group = par)) +
    geom_density(lwd = 0.3) +
    scale_x_continuous("") +
    scale_y_continuous("Density") +
    facet_grid(par ~ state, scales = "free_x") 
  
  p_5 <- ggplot(
    d_fig_2[par %like% "delta"], aes(x = value, group = par)) +
    geom_density(lwd = 0.3) +
    scale_x_continuous("") +
    scale_y_continuous("Density") +
    ggh4x::facet_grid2(par ~ state, scales = "free_x") 

  
  
  list(
    l_cfg = l_cfg,
    d_all = d_all,
    d_post_smry_1 = d_post_smry_1,
    d_post_smry_2 = d_post_smry_2,
    d_post_smry_3 = d_post_smry_3,
    d_post_smry_4 = d_post_smry_4,
    d_tran = d_tran,
    d_post_sop = d_post_sop,
    p_obs = p_1,
    p_cmp_1 = p_2,
    p_cmp_2 = p_3,
    p_dur = p_4 ,
    p_del = p_5
  )
  
}
```


## Example Trial 1


```{r, echo = F}
id_scen <- 1
id_ex <- 1
include_tran_m <- F
l_ex <- ex_trial_plots(id_scen, id_ex, include_tran_m)

```

In the example trial `r length(unique(l_ex$d_all$id))` were randomised.

### Trial Data

```{r}
#| label: fig-ex-1-obs
#| fig-cap: 'Summary of trial data: observed proportion in each state by visit day'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H


print(l_ex$p_obs)
```


### Transition Matrices

Transition matricies only reported if `include_tran_m` is set to `TRUE`, current value is `r include_tran_m`.

```{r, eval = include_tran_m}
#| label: tbl-ex-1-tran
#| tbl-cap: 'Transition matrices'

if(nrow(l_ex$d_tran) > 0){
  options(knitr.kable.NA = "-")
  kableExtra::kbl(
    l_ex$d_tran[day <= 14], 
    booktabs = TRUE, align = c("lllccc"), longtable = TRUE,
    linesep = c(rep("", 8), "\\addlinespace"),
    col.names=c("Day", "Treatment", "From", "none", "mild", "severe"), 
    digits = 3)  |>
    kableExtra::kable_styling(
      font_size = 8,
      latex_options = c("scale_down", "repeat_header")) |>
    kableExtra::column_spec(1:6, width = "2cm")
}


```


### SOPs

```{r}
#| label: fig-ex-1-sop-1
#| fig-cap: 'Summary of trial data vs model SOP'
#| fig-height: 4
#| fig-width: 6
#| out-height: 4in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H


print(l_ex$p_cmp_1)
```


### Soujourn Times

```{r}
#| label: fig-ex-1-soujourn
#| fig-cap: 'Posterior density of duration in each state at last analysis completed'
#| fig-height: 5
#| fig-width: 6
#| out-height: 5in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

print(l_ex$p_dur)
```



### Difference in Soujourn Times

```{r}
#| label: fig-ex-1-trt-eff
#| fig-cap: 'Posterior density of difference in duration in each state (trt - soc) at last analysis completed'
#| fig-height: 5
#| fig-width: 6
#| out-height: 5in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

print(l_ex$p_del)
```


### Trial Decisions

```{r}
#| label: tbl-ex-1-decision
#| tbl-cap: 'Trial decisions by analysis (based on difference in symptom free days)'

d_tbl_1 <- l_ex$d_post_smry_3[
  par %in% c("delta_def", "delta_dis") & state == "none", 
  .(ic, par, mu, lo, hi)]
d_tbl_2 <- l_ex$d_post_smry_4[, .(ic, rule, par, p)]
d_tbl_2 <- dcast(d_tbl_2, ic + par ~ rule, value.var = "p")

d_tbl <- base::merge(
  d_tbl_1,
  d_tbl_2, 
  by = c("ic", "par")
)
d_tbl[!is.na(mu), out := sprintf("%.2f (%.2f, %.2f)", mu, lo, hi)]
d_tbl <- d_tbl[, .(ic, par, out, ni, fut)]

options(knitr.kable.NA = "-")
kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llccc"), longtable = TRUE,
  linesep = c(rep("", 1), "\\addlinespace"),
  col.names=c("Analysis", "Par", "Mean (CrI)", "Pr(NI)", "Pr(fut)"), 
  digits = 3)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "1.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3, width = "2.5cm")|>
  kableExtra::column_spec(4:5, width = "2cm")
```

{{< pagebreak >}}


## Example Trial 2


```{r, echo = F}
id_scen <- 1
include_tran_m <- F
l_ex <- ex_trial_plots(id_scen, 3, include_tran_m)

```

In the example trial `r length(unique(l_ex$d_all$id))` were randomised.

### Trial Data

```{r}
#| label: fig-ex-2-obs
#| fig-cap: 'Summary of trial data: proportion in each state by day'
#| fig-height: 6
#| fig-width: 6
#| out-height: 6in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H


print(l_ex$p_obs)
```


### Transition Matrices

Transition matricies only reported if `include_tran_m` is set to `TRUE`, current value is `r include_tran_m`.

```{r, eval = include_tran_m}
#| label: tbl-ex-2-tran
#| tbl-cap: 'Transition matrices'

if(nrow(l_ex$d_tran) > 0){
  options(knitr.kable.NA = "-")
  kableExtra::kbl(
    l_ex$d_tran[day <= 14], 
    booktabs = TRUE, align = c("lllccc"), longtable = TRUE,
    linesep = c(rep("", 8), "\\addlinespace"),
    col.names=c("Day", "Treatment", "From", "none", "mild", "severe"), 
    digits = 3)  |>
    kableExtra::kable_styling(
      font_size = 8,
      latex_options = c("scale_down", "repeat_header")) |>
    kableExtra::column_spec(1:6, width = "2cm")
}

```

### SOPs


```{r}
#| label: fig-ex-2-sop-1
#| fig-cap: 'Summary of trial data vs model SOP'
#| fig-height: 4
#| fig-width: 6
#| out-height: 4in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H


print(l_ex$p_cmp_1)
```


### Soujourn Times

```{r}
#| label: fig-ex-2-soujourn
#| fig-cap: 'Posterior density of duration in each state at last analysis completed'
#| fig-height: 5
#| fig-width: 6
#| out-height: 5in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

print(l_ex$p_dur)
```

### Difference in Soujourn Times

```{r}
#| label: fig-ex-2-trt-eff
#| fig-cap: 'Posterior density of difference in duration in each state (trt - soc) at last analysis completed'
#| fig-height: 5
#| fig-width: 6
#| out-height: 5in
#| out-width: 6in
#| fig-dpi: 240
#| fig-align: "center"
#| fig-pos: H

print(l_ex$p_del)
```



### Trial Decisions

```{r}
#| label: tbl-ex-2-decision
#| tbl-cap: 'Trial decisions by analysis (based on difference in symptom free days)'

d_tbl_1 <- l_ex$d_post_smry_3[
  par %in% c("delta_def", "delta_dis") & state == "none", 
  .(ic, par, mu, lo, hi)]
d_tbl_2 <- l_ex$d_post_smry_4[, .(ic, rule, par, p)]
d_tbl_2 <- dcast(d_tbl_2, ic + par ~ rule, value.var = "p")

d_tbl <- base::merge(
  d_tbl_1,
  d_tbl_2, 
  by = c("ic", "par")
)
d_tbl[!is.na(mu), out := sprintf("%.2f (%.2f, %.2f)", mu, lo, hi)]
d_tbl <- d_tbl[, .(ic, par, out, ni, fut)]

options(knitr.kable.NA = "-")
kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llccc"), longtable = TRUE,
  linesep = c(rep("", 1), "\\addlinespace"),
  col.names=c("Analysis", "Par", "Mean (CrI)", "Pr(NI)", "Pr(fut)"), 
  digits = 3)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "1.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3, width = "2.5cm")|>
  kableExtra::column_spec(4:5, width = "2cm")
```


{{< pagebreak >}}

## Simulation Results

### Parameters 

@tbl-par-cmp-1 summarises the model parameter estimates, by scenario and analysis, averged over the simulations.

```{r, eval = T}
#| label: tbl-par-cmp-1
#| tbl-cap: 'Posterior parameter summary (model parameters)'

# l_spec_1$non_zero_pars

i <- 1
d_par <- data.table()
d_par_tru <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_tmp <- copy(l[[i]]$d_post_smry_1[par %in% l_spec_1$non_zero_pars])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate to impute the effects in those sims that had
  # early stopping
  d_tmp[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  d_smry <- d_tmp[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_par <- rbind(
    d_par,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
  
  d_tmp <- data.table(
    desc = l_cfg$desc, 
    par = l_spec_1$non_zero_pars,
    tru = c(
      l_cfg$alpha, 
      l_cfg$b_trt[-1], 
      l_cfg$b_prev[-1],
      l_cfg$b_time_1, l_cfg$b_time_2,
      l_cfg$b_prev_time[-1],
      l_cfg$b_trt_time[-1]
      )
    )
  
  d_par_tru <- rbind(
    d_par_tru,
    d_tmp
  )
}

# d_par_tru

# d_par[, stat := sprintf("%.3f \\n (%.3f, %.3f)", mu, lo, hi)]

d_par[, stat := sprintf("\\makecell[c]{%.3f\\\\(%.3f, %.3f)}", mu, lo, hi)]

# d_par[, stat := sprintf("%.3f\\newline(%.3f, %.3f)",mu, lo, hi)]

d_tbl <- dcast(d_par, desc + par ~ N, value.var = "stat")
d_tbl <- base::merge(
  d_tbl,
  d_par_tru,
  by = c("desc", "par"),
  all.x = T
)


d_tbl[, par := factor(par, levels = l_spec_1$non_zero_pars)]
d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl <- d_tbl[order(desc, par)]
d_tbl[, par := gsub("_", "\\\\_", par)]


k_out <- kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llcccc"), longtable = TRUE,
  # linesep = c(rep("", 11), "\\addlinespace"),
  linesep = "",
  col.names=c("Scenario", "Parameter", cumsum(l_cfg$N_pt), "True Value"), 
  digits = 3,
  escape = F)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3:5, width = "2cm") 

N_pars <- length(unique(d_tbl$par))
hline_seq_ix <- seq(N_pars, nrow(d_tbl), by = N_pars)
seq_ix <- seq_along(hline_seq_ix)[-length(hline_seq_ix)]

for(i in seq_ix){
  k_out <- k_out |>
    kableExtra::row_spec(hline_seq_ix[i], hline_after = TRUE)
}
k_out
```

{{< pagebreak >}}

```{r, eval = T}
#| label: tbl-par-cmp-2
#| tbl-cap: 'Posterior parameter summary (derived parameters - sojourn times)'

i <- 1
d_par <- data.table()
d_par_tru <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_tmp <- copy(l[[i]]$d_post_smry_3[par %in% c("soc", "def", "dis") & state == "none"])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate to impute the effects in those sims that had
  # early stopping
  d_tmp[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  d_smry <- d_tmp[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_par <- rbind(
    d_par,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
  
  d_tmp <- data.table(
    desc = l_cfg$desc, 
    par = l_spec_1$trt_lab,
    tru = c(
      l_cfg$dur_tru
      )
    )
  
  d_par_tru <- rbind(
    d_par_tru,
    d_tmp
  )
}


d_par[, stat := sprintf("\\makecell[c]{%.3f\\\\(%.3f, %.3f)}", mu, lo, hi)]


d_tbl <- dcast(d_par, desc + par ~ N, value.var = "stat")
d_tbl <- base::merge(
  d_tbl, 
  d_par_tru, 
  by = c("desc", "par"),
  all.x = T
)


d_tbl[, par := factor(par, levels = l_spec_1$trt_lab)]
d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl <- d_tbl[order(desc, par)]
d_tbl[, par := gsub("_", "\\\\_", par)]


k_out <- kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llcccc"), longtable = TRUE,
  linesep = c(rep("", 10), "\\addlinespace"),
  col.names=c("Scenario", "Parameter", cumsum(l_cfg$N_pt), "True value"), 
  digits = 3,
  escape = F)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3:5, width = "2.25cm") |>
  kableExtra::column_spec(6, width = "1.5cm")

N_pars <- length(unique(d_tbl$par))
hline_seq_ix <- seq(N_pars, nrow(d_tbl), by = N_pars)
seq_ix <- seq_along(hline_seq_ix)[-length(hline_seq_ix)]

for(i in seq_ix){
  k_out <- k_out |>
    kableExtra::row_spec(hline_seq_ix[i], hline_after = TRUE)
}
k_out
```

{{< pagebreak >}}

```{r, eval = T}
#| label: tbl-par-cmp-3
#| tbl-cap: 'Posterior parameter summary (derived parameters - difference in sojourn times)'

i <- 1
d_par <- data.table()
d_par_tru <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_tmp <- copy(l[[i]]$d_post_smry_3[par %in% c("delta_def", "delta_dis") & state == "none"])
  
  l_cfg <- copy(l[[i]]$l_spec)
  
  # extrapolate to impute the effects in those sims that had
  # early stopping
  d_tmp[, mu := nafill(mu, type = "locf"), keyby = .(sim, par)]
  
  d_smry <- d_tmp[, .(
    mu = mean(mu),
    lo = quantile(mu, prob = 0.025),
    hi = quantile(mu, prob = 0.975)
    ), keyby = .(ic, par)]
  
  d_smry[, N := cumsum(l_cfg$N_pt)[ic]]
    
  d_par <- rbind(
    d_par,
    cbind(scenario = i, desc = l_cfg$desc, d_smry)
  )
  
  d_tmp <- data.table(
    desc = l_cfg$desc, 
    par = l_spec_1$delta_lab,
    tru = c(
      l_cfg$delta_dur_tru
      )
    )
  
  d_par_tru <- rbind(
    d_par_tru,
    d_tmp
  )
}


d_par[, stat := sprintf("\\makecell[c]{%.3f\\\\(%.3f, %.3f)}", mu, lo, hi)]


d_tbl <- dcast(d_par, desc + par ~ N, value.var = "stat")
d_tbl <- base::merge(
  d_tbl, 
  d_par_tru, 
  by = c("desc", "par"),
  all.x = T
)


d_tbl[, par := factor(par, levels = c("delta_def", "delta_dis"))]
d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl <- d_tbl[order(desc, par)]
d_tbl[, par := gsub("_", "\\\\_", par)]


k_out <- kableExtra::kbl(
  d_tbl, 
  booktabs = TRUE, align = c("llcccc"), longtable = TRUE,
  linesep = c(rep("", 10), "\\addlinespace"),
  col.names=c("Scenario", "Parameter", cumsum(l_cfg$N_pt), "True value"), 
  digits = 3,
  escape = F)  |>
  kableExtra::kable_styling(
    font_size = 8,
    latex_options = c("scale_down", "repeat_header")) |>
  kableExtra::column_spec(1, width = "3.5cm")|>
  kableExtra::column_spec(2, width = "2cm")|>
  kableExtra::column_spec(3:5, width = "2cm") |>
  kableExtra::column_spec(6, width = "1.5cm")



N_pars <- length(unique(d_tbl$par))
hline_seq_ix <- seq(N_pars, nrow(d_tbl), by = N_pars)
seq_ix <- seq_along(hline_seq_ix)[-length(hline_seq_ix)]

for(i in seq_ix){
  k_out <- k_out |>
    kableExtra::row_spec(hline_seq_ix[i], hline_after = TRUE)
}
k_out
```

### Decision Probability

@tbl-dec-pr-1 shows the cumulative probability of the non-inferiority and futility decisions based on the difference in the exacerbation sojourn times.

```{r, echo = F, eval = T}
#| label: tbl-dec-pr-1
#| tbl-cap: 'Cumulative probability of each decision type'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H


i <- 1
d_dec <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_post_smry_4[par %in% c("delta_def", "delta_dis")])
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  
  # number of enrolments at each interim (interim sample size sequence)
  d_N <- data.table(ic = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  d_dec_1[, cdec := as.integer(cumsum(dec)>0), keyby = .(sim, rule, par)]
  d_dec_1[, cdec := nafill(cdec, type = "locf"), keyby = .(sim, rule, par)]
  d_dec_1[, cdec := as.logical(cdec)]
  
  d_dec_1 <- base::merge(d_dec_1, d_N, by = "ic")
  
  # cumulative proportion for which each decision quantity has been met by 
  # analysis and domain
  d_dec_1 <- d_dec_1[, .(pr_val = mean(cdec)), keyby = .(ic, N, rule, par)]
  
  d_dec <- rbind(
    d_dec,
    cbind(scenario = i, desc = l_cfg$desc, d_dec_1)
  )

}

d_tbl <- dcast(
  d_dec, desc + rule  + par ~ N, value.var = "pr_val")

d_tbl[, desc := factor(desc, levels = desc_labs)]
d_tbl[, rule := factor(rule, levels = c("ni", "fut"), labels = c("NI", "Fut"))]
d_tbl <- d_tbl[order(desc, rule)]


kableExtra::kbl(
  d_tbl, 
  escape = TRUE, booktabs = TRUE, align = c("lllccc"),
  linesep = c("", "", "", "\\addlinespace"),
  col.names=c("Scenario", "Rule", "Parameter", cumsum(l_cfg$N_pt)), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 8) |>
  kableExtra::column_spec(1, width = "5cm")|>
  kableExtra::column_spec(2:6, width = "1.5cm")



```





{{< pagebreak >}}

### Sample Size

Expected number of enrolments on each treatment arm under each scenario.

```{r, echo = F, eval = F}
#| label: tbl-samp-size
#| tbl-cap: 'Mean sample size by scenario and treatment group'
#| column: page
#| out-width: 100%
#| tbl-cap-location: bottom
#| tbl-pos: H

i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
   
  d_all <- copy(l[[i]]$d_all[evt == 1])
  
  d_smry <- d_all[, .N, keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_smry[, .(mu_N = mean(N), sd_N = sd(N)), keyby = trt]
    )
  )
  
}
d_N_by_rand[, out := sprintf("%.0f (%.0f)", mu_N, sd_N)]
d_tbl <- dcast(d_N_by_rand, scenario + desc ~ trt, value.var = "out")
d_tbl <- d_tbl[order(scenario)]


kableExtra::kbl(
  d_tbl[, .SD, .SDcols = -c("scenario")], 
  escape = TRUE, booktabs = TRUE, align = c("lrrr"),
  col.names=c("Scenario", "def", "dis", "soc"), 
  digits = 2)  |>
  kableExtra::kable_styling(font_size = 8) 


```



{{< pagebreak >}}

# References {.unnumbered}

<!-- Needs to have a citation for this to work otherwise you will get the old \end{CSLReferences} error  -->

::: {#refs}
:::

{{< pagebreak >}}




