---
title: BEAT CF MARS
subtitle: Simulation Report
description: |
    Three-arm Bayesian adaptive trial nested within BEAT-CF platform evaluating antibiotic-sparing strategies for managing pulmonary exacerbations
date: last-modified
date-format: "D MMMM YYYY"
author: 
  - name: Mark Jones
    id: mj
    email: mark.jones1@sydney.edu.au
version: 0.1
sponsor: "University of Sydney, NSW, Australia"
protocol-number: todo
registration: todo
hrec: todo
ci1: Tom Snelling
editor: source
bibliography: ../etc/refs.bib
csl: ../etc/elsevier-harvard.csl
# number-sections required otherwise section refs will not render 
number-sections: true
toc: true
toc-depth: 3
format:
  pdf: 
    pdf-engine: xelatex
    keep-tex: true
    documentclass: scrreprt
    papersize: a4
    fontsize: 12pt
    mainfont: Libertinus Serif
    sansfont: Libertinus Sans
    monofont: Libertinus Mono
    mathfont: Libertinus Math
    linestretch: 1.25
    template-partials: 
      - "../_extensions/partials/before-body.tex"
    include-in-header:
      text: |
       \usepackage{physics}
       \setkomafont{chapter}{\fontsize{16}{18}\selectfont}
       \setkomafont{section}{\fontsize{14}{16}\selectfont}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
)
```

```{r}
#| echo: false

# uml digs
suppressPackageStartupMessages(library(nomnoml))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(qs))
suppressPackageStartupMessages(library(git2r))
suppressPackageStartupMessages(suppressWarnings(library(gt)))
suppressPackageStartupMessages(library(ggh4x))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(patchwork))
suppressPackageStartupMessages(library(pbapply))


toks <- unlist(tstrsplit(getwd(), "/")) 
if(toks[length(toks)] == "beatcf-mars-sim"){
  prefix_cfg <- "./etc/sim12/"
  prefix_stan <- "./stan"
  prefix_fig <- "./fig"
  prefix_data <- "./data"
  prefix_r <- "./r"
} else {
  prefix_cfg <- "../etc/sim12/"
  prefix_stan <- "../stan"
  prefix_fig <- "../fig"
  prefix_data <- "../data"
  prefix_r <- "../r"
}


source(paste0(prefix_r, "/data.R"))
source(paste0(prefix_r, "/data-sim12.R"))


# Each input file corresponds to the results from a single simulation
# scenario/configuration.
# Load all the files into a single list.

# files of interest
sim_lab <- "sim12-02"

flist <- list.files(paste0(prefix_data, "/", sim_lab), pattern = "sim12")
toks <- list()
l <- list()
i <- 1

for(i in 1:length(flist)){

  l[[i]] <- qs::qread(file.path(paste0(prefix_data, "/", sim_lab), flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

N_sims <- l[[1]]$cfg$nsim



# temp to replace
# l_spec <- get_prototype_cfg()
l_spec_1 <- l[[1]]$l_spec



```

{{< pagebreak >}}

::: summary
|     |        |
|:----|:------------|
|Study title:  |  BEAT-CF MARS |
|Reference intervention: |  Standard airway clearance therapy at least twice per day for 14 days, plus 14 days of oral antibiotics commencing as soon as practicable (and ≥ 72 hours and < 7 days) after onset of new or acute worsening of cough. |   
|Intervention: | (1) Deferred antibiotic strategy (2) Early discontinuation antibiotic strategy. |   
|Outcome: | ppFev1 at 12 months |   
|Study design:  |   Bayesian adaptive trial with early stopping rules | 
|Sponsor:  |    University of Sydney, NSW, Australia | 
Protocol: |  todo |
|Registration:  |    todo | 
|HREC:  |   todo | 
|Study date of first consent:  |   todo | 
|Principal coordinating investigators:  |   Tom Snelling | 
:::

<!-- 
Note that the above relies on the pandoc extension implemented in the lua file 
in the etc directory. It additionally relies on the presence of a custom style
in word called study summary. It will currently only work for word (because I
cannot be bothered to implement it in anything else at the moment).
-->


{{< pagebreak >}}

# Version history {.unlisted .unnumbered}

| Version    |   Date     | Change    |   Reason     |
|:----|:------------|:----|:------------|
| 0.1 | 2025-11-12 | First version | N/A |


{{< pagebreak >}}



# Introduction

This report documents the simulation approach and results for the operating characteristics for the BEAT-CF MARS study.
The report is an operational document that will be updated, as necessary, over the course of the study.
It should be read in conjunction with the relevant versions of the protocol and statistical analysis plan.

We provide the data generation assumptions, modelling approaches, scenarios and results that were used to explore the design.

These results are based on simulation `r sim_lab` with `r N_sims` simulated trials run per scenario.

The key configuration parameters that persist across all scenarios are shown below:

+ Interim analyses: `r cumsum(l_spec_1$N)`
+ Futility evaluation at: `r cumsum(l_spec_1$N)`
+ Non-inferiority evaluation at: `r cumsum(l_spec_1$N)`
+ Enrolment per day after ramp up: `r l_spec_1$pt_per_day`
+ Ramp up period (days): `r l_spec_1$ramp_up_days`
+ Standard deviation on baseline ppFEV1: `r l_spec_1$sigma_mu0`
+ Standard deviation on ppFEV1: `r l_spec_1$sigma`
+ Correlation between repeat measures: `r l_spec_1$rho`
+ Reference value for non-inferiority: `r l_spec_1$delta$ni`
+ Evidential threshold for non-inferiority (probability): `r l_spec_1$thresh$ni`
+ Reference value for futility: `r l_spec_1$delta$fut`
+ Evidential threshold for futility (probability): `r l_spec_1$thresh$fut`



 

# Design overview





# Results


## Probability of triggering decisions

@tbl-cprob-decision provides the cumulative probability of non-inferiority by scenario with the probability of declaring futility in parentheses.

```{r, echo = F, eval = T}
#| label: cum_prob_dec
#| code-summary: Cumulative probability of each decision type

# Cumulative probability of decisions:

# Traverse the list of simulation results and for each one summarise the 
# cumulative probability of each decision type.

i <- 1
d_cprob_dec <- data.table()

# For each scenario that was simulated
for(i in 1:length(l)){
  
  # extract the decision matrix - sim, analysis, quantity, domain level decision
  d_dec_1 <- copy(l[[i]]$d_pr_dec[trt %in% c("delay", "defer")])
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  
  # number of enrolments at each interim (interim sample size sequence)
  d_N <- data.table(ic = seq_along(l_cfg$N_pt), N = cumsum(l_cfg$N_pt))
  
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  d_dec_1[, cdec := as.integer(cumsum(dec)>0), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := nafill(cdec, type = "locf"), keyby = .(sim, rule, trt)]
  d_dec_1[, cdec := as.logical(cdec)]
  
  d_dec_1 <- base::merge(d_dec_1, d_N, by = "ic")
  
  # cumulative proportion for which each decision quantity has been met by 
  # analysis and domain
  d_dec_1 <- d_dec_1[, .(pr_val = mean(cdec)), keyby = .(ic, N, rule, trt)]
  
  
  d_cprob_dec <- rbind(
    d_cprob_dec,
    cbind(scenario = i, desc = l_cfg$desc, d_dec_1)
  )

}

```





```{r, eval = T}
#| echo: FALSE
#| label: tbl-cprob-decision
#| tbl-cap: 'Cumulative probability of decision for each enrolment cohort'
#| tbl-pos: H

d_tbl_1_cur <- copy(d_cprob_dec)
d_tbl_1_cur <- dcast(
  d_tbl_1_cur, scenario + desc + trt ~ rule + N, value.var = "pr_val")

# d_b_trt <- data.table()
# 
# for(i in 1:length(l)){
#   l_cfg <- copy(l[[i]]$cfg)
#   
#   # main effect
#   b_T <- unlist(l_cfg$b_T)
#   # time interaction
#   b_tT <- l_cfg$b_tT
#   # cumulative effects
#   b_trt <- c(
#     b_T[1] + sum(b_tT[1, ]),
#     b_T[2] + sum(b_tT[2, ])
#   )
#   
#   d_tmp <- data.table(
#     scenario = i, 
#     par = c("Deferred", "Discontinued"), 
#     delta = b_trt
#   )
#   d_b_trt <- rbind(d_b_trt, d_tmp)
# }

# d_tbl_1_cur <- base::merge(d_tbl_1_cur, d_b_trt, by = c("scenario", "par"))

# setcolorder(d_tbl_1_cur, c("scenario", "desc", "par", "delta"))
# setorderv(d_tbl_1_cur, cols = c("scenario"))

d_tbl_1_cur <- d_tbl_1_cur[, .SD, .SDcols = !c("scenario")]

g_tbl <- d_tbl_1_cur |> 
  gt(groupname_col = "desc") |> 
  gt::text_transform(
    locations = cells_row_groups(),
    fn = function(x) {
      lapply(x, function(x) {
        gt::md(paste0("*", x, "*"))
      })
    }
  ) |>
  cols_align(
    columns = 1:2,
    align = "left"
  )  |> 
  cols_align(
    columns = 3:ncol(d_tbl_1_cur),
    align = "center"
  )  |> 
  cols_merge(
    columns = c("ni_400", "fut_400"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |> 
  cols_merge(
    columns = c("ni_500", "fut_500"
                ),
    pattern = "<<{1}>><< ({2})>>"
  )   |>
  cols_merge(
    columns = c("ni_600", "fut_600"
                ),
    pattern = "<<{1}>><< ({2})>>"
  )   |> 
  cols_width(
    starts_with("ni") ~ px(100)
  ) |>
  tab_spanner(
    label = md("Participants having reached primary endpoint"),
    columns = 4:ncol(d_tbl_1_cur)
  )  |>
  cols_label(
    # delta = "Diff at 12 month vs SoC",
    trt = "",
    ni_400 = html("400"),
    ni_500 = html("500"),
    ni_600 = html("600")
  ) |>
  tab_options(
    table.font.size = pct(55),
    latex.use_longtable = TRUE,
    latex.header_repeat = TRUE 
  ) |>
  fmt_number(decimals = 2, drop_trailing_zeros = F)

g_tbl
```


{{< pagebreak >}}

## Sample size

@tbl-n-1 shows the expected number of participants on each treatment arm under each scenario.

```{r, echo = F, eval = T}
#| label: sample_size_1
#| code-summary: Compute expected sample size


i <- 1
d_N_by_rand <- data.table()

for(i in 1:length(l)){
  
  # config for scenario
  l_cfg <- copy(l[[i]]$l_spec)
  # l_cfg$desc
  
  N_analysis <- length(l_cfg$N_pt)
  N_trt <- 3
  # 
  d_grid <- CJ(
    sim = 1:l_cfg$nsim,
    ic = 1:N_analysis,
    trt = 1:N_trt
  )
  d_grid$trt <- l[[i]]$l_spec$trt_lab[d_grid$trt]
    
  # pick up the baseline obs only, i.e. when the number enrolled
  d_all <- copy(l[[i]]$d_all[state == "E"])
  d_analy_set <- copy(l[[i]]$d_analys_sets)
  d_analy_set <- d_analy_set[!is.na(t0_last), ]
  d_analy_set[, ic := 1]
  d_analy_set[, ic := cumsum(ic), by = sim]
  
  setkey(d_all, sim, t0)
  setkey(d_analy_set,  sim, t0_last)

  # Rolling join: find first cutoff_date >= enrol_date
  d_all[
    d_analy_set,
    on = .(sim, t0 = t0_last),
    roll = -Inf,
    ic := i.ic
  ]
  
  d_all[, ic := nafill(ic, type = "nocb"), by = sim]
  
  # how many randomisations per interim and trt arm (pt may have more than 
  # one randomisation over their followup)
  d_smry <- base::merge(
    d_grid,
    d_all[, .(.N), keyby = .(sim, ic, trt)],
    by = c("sim", "ic", "trt"),
    all.x = T
  )
  
  # some arms may stop after interim 1 and so there would be no count for 
  # subsequent interims
  d_smry[is.na(N), N := 0]
  d_smry[, N_cmtv := cumsum(N), keyby = .(sim, trt)]
  
  d_N_by_rand <- rbind(
    d_N_by_rand,
    cbind(
      scenario = i, desc = l_cfg$desc, 
      d_smry
    )
  )
  
}
```




```{r, eval = T}
#| echo: FALSE
#| label: tbl-n-1
#| tbl-cap: ''
#| tbl-pos: H

# DONT assume that all arms ran to the final interim. The result is just an
# artefact of imputation. Nevertheless it gives the true sample size by arm.
d_N_trt <- d_N_by_rand[, .SD[.N], keyby = .(scenario, desc, sim, trt)]
d_N_tot <- d_N_trt[, .(N_cmtv = sum(N_cmtv)), keyby = .(scenario, desc, sim) ]

# add in total sample size
d_tbl_1_cur <- rbind(
  d_N_trt[, .(N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc, trt)],
  d_N_tot[, .(trt = "all", N_mu = mean(N_cmtv), N_sd = sd(N_cmtv)), keyby = .(scenario, desc)]
)


# we are only interested in Deferred vs soc and Discontinued vs soc
d_b_trt <- data.table()

# for(i in 1:length(l)){
#   l_cfg <- copy(l[[i]]$cfg)
#   
#   # main effect
#   b_T <- unlist(l_cfg$b_T)
#   # time interaction
#   b_tT <- l_cfg$b_tT
#   # cumulative effects
#   b_trt <- c(
#     b_T[1] + sum(b_tT[1, ]),
#     b_T[2] + sum(b_tT[2, ])
#   )
#   
#   d_tmp <- data.table(
#     scenario = i, 
#     trt = c("Deferred", "Discontinued"), 
#     delta = c(b_trt[1], b_trt[2]))
#   d_b_trt <- rbind(d_b_trt, d_tmp)
# }

# d_tbl_1_cur <- base::merge(
#   d_tbl_1_cur, 
#   d_b_trt, 
#   by = c("scenario", "trt"),
#   all.x = T)
# 
d_tbl_1_cur <- dcast(
  d_tbl_1_cur,
  scenario + desc ~ trt ,
  value.var = list("N_mu", "N_sd"))

# d_tbl_1_cur[, `:=`(delta_SoC = NULL, delta_all = NULL)]

# setcolorder(d_tbl_1_cur, c("scenario", "desc", "delta_Deferred", "delta_Discontinued"))

# d_tbl_1_cur <- d_tbl_1_cur[order(scenario, desc)]
d_tbl_1_cur <- d_tbl_1_cur[, .SD, .SDcols = !c("scenario")]



g_tbl <- d_tbl_1_cur |> 
  gt(groupname_col = "desc")  |> 
  gt::text_transform(
    locations = cells_row_groups(),
    fn = function(x) {
      lapply(x, function(x) {
        gt::md(paste0("*", x, "*"))
      })
    }
  ) |>
  fmt_number(
    columns = starts_with("N_"),
    decimals = 0, drop_trailing_zeros = TRUE) |>
  # fmt_number(
  #   columns = starts_with("delta"),
  #   decimals = 2, drop_trailing_zeros = TRUE) |> 
  # tab_spanner(
  #   label = md("Diff at 12 month vs SoC"),
  #   columns = starts_with("delta"),
  # ) |>
  tab_spanner(
    label = md("Sample size, mean (sd)"),
    columns =  starts_with("N_mu")
  ) |>
  # cols_align(
  #   columns = starts_with("delta"),
  #   align = "center"
  # ) |>
  cols_align(
    columns = starts_with("N_mu"),
    align = "center"
  )  |>
  cols_merge(
    columns = c("N_mu_defer", "N_sd_defer"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_delay", "N_sd_delay"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_soc", "N_sd_soc"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_merge(
    columns = c("N_mu_all", "N_sd_all"
                ),
    pattern = "<<{1}>><< ({2})>>"
  ) |>
  cols_label(
    # delta_Deferred = "defer",
    # delta_Discontinued = "delay",
    N_mu_defer = "defer",
    N_mu_delay = "delay",
    N_mu_soc = "soc",
    N_mu_all = "Total"
  ) |>
  tab_options(
    # table.width = pct(100),
    table.font.size = pct(55)
  ) 


g_tbl
```